### 2018-10-11 BF_2D_notebook_01.Rmd
added BF_2D  notebook

---
title: "Basis functions approximations for reduced-rank Gaussian process regression in high-dimensinal space"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(viridis)
library(rstan)
library(rgl)
library(RColorBrewer)
library(brms)
library(rstanarm)
library(geoR)
library(lattice)
library(aqfig)
library(fields)
library(colorRamps)
```

<!-- Gaussian Processes are flexible non-parametric stochastic models for $D$-dimensional functions. However, a Gaussian process model suffer from observational dimensionality $(N)$, where the inversion of the covariance matrix requires of $O(N^3)$ computational expenses.  -->

<!-- This notebook deals with the implementation of an approximate eigendecomposition of the covariance function of a D-dimensional Gaussian process. The method is based on interpreting the covariance function as the kernel of a pseudo-differential operator and approximating it using Hilbert space methods. This results in a reduced-rank approximation for the covariance function (Solin and Särkkä, 2018). Comparison of this approximated framework with the full Gaussian process and a splines-based model is carried out.  -->

### D-dimensional Gaussian process model

(Formulation of the D-dimensional GP model is missing)

### Approximate eigendecomposition of the covariance function of a D-dimensional Gaussian process

(Formulation of the D-dimensional Basis function aproximation model is missing)

### Application to simulated data

Simulating data from the following 2D function $f(x_1,x_2)$ and contaminate it with Gaussian noise $e \sim N(0,\sigma^2)$.

$$
f(x_1,x_2) = \frac{1}{100} \cdot (x_1^3 - x_2^3) + \frac{1}{10} \cdot x_1 \cdot x_2
$$
$$
y(x_1,x_2) = f(x_1,x_2) + e
$$

```{r echo=TRUE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# color palettes
col1 <- brewer.pal(11,"BrBG")[11:1] 
col2 <- cm.colors(11, alpha = 0.5)
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(15), colorRampPalette(c("white", "red"), bias=10)(15)[2:15])

# number of simulated observations and true function
nobs <- 120  
f <- function(x1,x2) 1/100*(x1^3 - x2^3) + 1/10*x1*x2

# grid input values
x_grid <- pred_grid(coords= c(-5,5), y.coords= c(-5,5), by= 0.3)

# breaks for colors
par(mfrow=c(1,2))
brk <- seq(min(f(x_grid[,1],x_grid[,2])),max(f(x_grid[,1],x_grid[,2])), length.out= 12)

# true function surface evaluated in the grid input values
xx <- unique(x_grid[,1])
image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, main="true function surface f and simulated locations", xlab="x1", ylab="x2", cex.main=0.7)

# training input values
set.seed(10)
x1 <- runif(nobs,-5,5)  
x2 <- runif(nobs,-5,5)

# simulated training observations
y <-  f(x1,x2) + rnorm(length(x1), 0, 0.3)

points(x1, x2, pch=1, cex=0.7)
plot(f(x1, x2), y, col=4, cex=0.6, main="true function values f vs noisy simulations y", xlab="f", ylab="y", cex.main=0.7); abline(a=0, b=1, col="gray")

# noisy observations in the grid input locations
y_grid <- f(x_grid[,1],x_grid[,2]) + rnorm(length(dim(x_grid)[1]), 0, 0.2)

# total (training and grid) input values
x <- rbind(cbind(x1,x2),as.matrix(x_grid))

# number of training, grid and total observations
N1 <- length(x1)
N_grid <- dim(x_grid)[1]
N <- N1 + N_grid
```

### 2D Gaussian process model

Standata.

```{r}
D <- 2  # number of dimensions

standata_gp12<- list(D= D, x= cbind(x1,x2), y= y, N1= N1, N_grid= N_grid, x_grid= x_grid, y_grid= y_grid)
```

Stancode.

(The code of the D-dimensional GP model is missing)

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_gp12.rData")

gp12_f <- summary(aj_gp12, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y <- summary(aj_gp12, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_f_grid <- summary(aj_gp12, pars = c("f_grid"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y_grid <- summary(aj_gp12, pars = c("y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_elpd_grid <- summary(aj_gp12, pars= c("log_y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary
gp12_time <- sum(get_elapsed_time(aj_gp12)[2])/(aj_gp12@sim$iter - aj_gp12@sim$warmup)
```

Posteriors plots.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(0, 0, 2, 1.5))

# training
plot(f(x1,x2), gp12_f[,1], col=4, cex=0.6, xlab="f", ylab="gp_f", main="true values f vs posterior means gp_f", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(a=0, b=1, col="gray")
plot(f(x1,x2), f(x1,x2) - gp12_f[,1], xlab="f", ylab="f - gp_f", main="posterior residuals (f - gp_f)", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=0, col="gray")

title(main= "GP model", outer= TRUE)

# predicting
plot(f(x_grid[,1],x_grid[,2]), gp12_f_grid[,1], col=4, cex=0.6, xlab="f", ylab="gp_f", main="true values f vs posterior means gp_f", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(a=0, b=1, col="gray")
plot(f(x_grid[,1],x_grid[,2]), f(x_grid[,1],x_grid[,2]) - gp12_f_grid[,1], xlab="f", ylab="f - gp12_f", main="posterior residuals (f - gp_f)", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=0, col="gray")
plot(f(x_grid[,1],x_grid[,2]), gp12_elpd_grid[,1], col=2, cex=0.8, xlab="f", ylab="ELPD", mgp= c(1.7, 0.5, 0), cex.main=0.9); abline(h=mean(gp12_elpd_grid[,1]), col="red", lty=2)

# histogram residuals
gp12_error_grid <- matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)) - gp12_f_grid[,1]

hist(gp12_error_grid, mgp= c(1.7, 0.5, 0), xlab="f - gp_f", ylab="", main="posterior residuals (f - gp_f)", cex.main=0.9)

#images
image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, main="true surface f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.9)

image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col= col1, breaks= brk , main="posterior mean surface gp_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.9)

brk1 <- seq(-max(abs(gp12_error_grid)),max(abs(gp12_error_grid)), length.out= 30)

image.plot(unique(x_grid[,1]), unique(x_grid[,1]), gp12_error_grid, col= col3, breaks= brk1, main="posterior residuals surface", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, cex.main=0.9)
points(x1, x2, pch=1, cex=0.6)
```

### 2D Basis functions model

Standata.

```{r}
M <- c(2,4,6,10,15,20,30)  #number of basis functions
D <- 2  #number of dimensions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2)
		}
	}
}

standat_bf12 <- list()
for(j in 1:length(M)){  #number of basis functions
  	standat_bf12[[j]] <- list(D= D, M= M[j], M_nD= M[j]^D, L= c(5/2*max(x1),5/2*max(x2)), x= rbind(cbind(x1,x2),as.matrix(x_grid)), y= c(y,y_grid), N1= N1, indices= indices[[j]], N_grid= N_grid)
}
```

Stancode.

(The code of the D-dimensional Basis function aproximation model is missing)

Load and store the outcomes of the different Basis functions models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_bf12.rData")

bf12_f <- list()
bf12_y <- list()
bf12_elpd <- list()
bf12_time <- list()
for(j in 1:length(M)){  #number of basis functions
	bf12_f[[j]] <- summary(aj_bf12[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_y[[j]] <- summary(aj_bf12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_elpd[[j]] <- summary(aj_bf12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_time[[j]] <- sum(get_elapsed_time(aj_bf12[[j]])[2])/(aj_bf12[[j]]@sim$iter - aj_bf12[[j]]@sim$warmup)
}
```

Posteriors plots as function of the number of basis functions.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
for(j in 1:length(M)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp12_f[,1], bf12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="gp_f ", ylab="bf_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f[,1], gp12_f[,1] - bf12_f[[j]][1:N1,1], xlab="gp_f ", ylab="gp_f - bf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - bf_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("BF model with M =",M[j]), outer= TRUE)

	# predicting
	plot(gp12_f_grid[,1], bf12_f[[j]][-(1:N1),1], col=4, cex=0.6, xlab="gp_f ", ylab="bf_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f_grid[,1], gp12_f_grid[,1] - bf12_f[[j]][-(1:N1),1], xlab="gp_f ", ylab="gp_f - bf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - bf_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp12_f_grid[,1], bf12_elpd[[j]][-(1:N1),1], col=2, cex=0.8, xlab="gp_f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(bf12_elpd[[j]][-(1:N1),1]), col="red", lty=2)

	# histogram residuals
	bf12_error <- matrix(gp12_f_grid[,1],length(xx),length(xx)) - bf12_f[[j]][-(1:N1),1]

	hist(bf12_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - bf_f", ylab="", main="posterior residuals (gp_f - bf_f)", cex.main=0.8)

	# images
	image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, main="GP posterior means surface gp_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.8)

	image(xx, xx, matrix(bf12_f[[j]][-(1:N1),1],length(xx),length(xx)), col= col1, breaks= brk, main="BF posterior means surface bf_f", xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), cex.main=0.8)

	brk1 <- seq(-max(abs(bf12_error)),max(abs(bf12_error)), length.out= 30)

	image.plot(xx, xx, bf12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.5)
}
```

### 2D Additive Gaussian process model

Standata.

```{r}
standat_AGP <- list(D= 2, x= cbind(x1,x2), y= y, N1= N1, N_grid= N_grid, x_grid= x_grid, y_grid= y_grid)
```

Stancode.

(The code of the D-dimensional additive GP model is missing)

Load and store the outcomes of the additive GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_gp1gp2.rData")

gp1gp2_f <- summary(aj_gp1gp2, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_y <- summary(aj_gp1gp2, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_f_grid <- summary(aj_gp1gp2, pars = c("f_grid"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_y_grid <- summary(aj_gp1gp2, pars = c("y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp1gp2_elpd_grid <- summary(aj_gp1gp2, pars= c("log_y_grid_predict"), probs = c(0.025, 0.5, 0.975), digits_summary= 4)$summary
gp1gp2_time <- sum(get_elapsed_time(aj_gp1gp2)[2])/(aj_gp1gp2@sim$iter - aj_gp1gp2@sim$warmup)
```

Posterior plots.

```{r echo=TRUE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(0, 0, 1.5, 1.5))

# training
plot(gp12_f[,1], gp1gp2_f[1:N1,1], col=4, cex=0.6, xlab="gp_f ", ylab="agp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs AGP posterior means agp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
plot(gp12_f[,1], gp12_f[,1] - gp1gp2_f[1:N1,1], xlab="gp12_f ", ylab="gp12_f - agp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - agp_f)", cex.main=0.8); abline(h=0, col="gray")

title(main= paste("Additive GP model"), outer= TRUE)

# predicting
plot(gp12_f_grid[,1], gp1gp2_f_grid[,1], col=4, cex=0.6, xlab="gp_f ", ylab="agp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs AGP posterior means agp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
plot(gp12_f_grid[,1], gp12_f_grid[,1] - gp1gp2_f_grid[,1], xlab="gp_f ", ylab="gp_f - agp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - agp_f)", cex.main=0.8); abline(h=0, col="gray")
plot(gp12_f_grid[,1], gp1gp2_elpd_grid[,1], col=2, cex=0.8, xlab="gp_f ", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(gp1gp2_elpd_grid[,1]), col="red", lty=2)

# histogram residuals
gp1gp2_error <- matrix(gp12_f_grid[,1],length(xx),length(xx)) - gp1gp2_f_grid[,1]

hist(gp1gp2_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - agp_f", ylab="", main="posterior residuals (gp_f - agp_f)", cex.main=0.8)

# images
image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="GP posterior means surface gp_f", cex.main=0.8)

image(xx, xx, matrix(gp1gp2_f_grid[,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="AGP posterior means surface agp_f", cex.main=0.8)

brk1 <- seq(-max(abs(gp1gp2_error)),max(abs(gp1gp2_error)), length.out= 30)

image.plot(xx, xx, gp1gp2_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
points(x1, x2, pch=1, cex=0.5)
```

### 2D Additive Basis functions model

Standata.

```{r}
M <- c(2,4,6,10,15,20,30,40)  # number of basis functions
D <- 2  # number of dimensions

standat_bf1bf2 <- list()
for(j in 1:length(M)){
	standat_bf1bf2[[j]] <- list(D= D, M= M[j], L= c(5/2*max(x1), 5/2*max(x2)), x= rbind(cbind(x1,x2),as.matrix(x_grid)), y= c(y,y_grid), N1= N1, N_grid= N_grid)
}
```

Stancode.

(The code of the D-dimensional additive Basis function aproximation model is missing)

Load and store the outcomes of the different additive BF models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_bf1bf2.rData")

bf1bf2_f <- list()
bf1bf2_y <- list()
bf1bf2_elpd <- list()
bf1bf2_time <- list()
for(j in 1:length(M)){  # number of Basis functions
	bf1bf2_f[[j]] <- summary(aj_bf1bf2[[j]], pars = c("F"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_y[[j]] <- summary(aj_bf1bf2[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_elpd[[j]] <- summary(aj_bf1bf2[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf1bf2_time[[j]] <- sum(get_elapsed_time(aj_bf1bf2[[j]])[2])/(aj_bf1bf2[[j]]@sim$iter - aj_bf1bf2[[j]]@sim$warmup)
}
```

Posterior plots as function of the number of basis functions.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
for(j in 1:length(M)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp1gp2_f[,1], bf1bf2_f[[j]][1:N1,1], col=4, cex=0.6, xlab="agp_f ", ylab="abf_f", mgp= c(1.7, 0.5, 0), main="AGP posterior means agp_f vs ABF posterior means abf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp1gp2_f[,1], gp1gp2_f[,1] - bf1bf2_f[[j]][1:N1,1], xlab="agp_f ", ylab="agp_f - abf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (agp_f - abf_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("Additive BF model with M =",M[j]), outer= TRUE)

	# predicting
	plot(gp1gp2_f_grid[,1], bf1bf2_f[[j]][-(1:N1),1], col=4, cex=0.6, xlab="agp_f ", ylab="abf_f", mgp= c(1.7, 0.5, 0), main="AGP posterior means agp_f vs ABF posterior means abf_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp1gp2_f_grid[,1], gp1gp2_f_grid[,1] - bf1bf2_f[[j]][-(1:N1),1], xlab="agp_f ", ylab="agp_f - abf_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (agp_f - abf_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp1gp2_f_grid[,1], bf1bf2_elpd[[j]][-(1:N1),1], col=2, cex=0.8, xlab="agp_f ", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(bf1bf2_elpd[[j]][,1]), col="red", lty=2)

	# histogram residuals
	bf12_error <- matrix(gp1gp2_f_grid[,1],length(xx),length(xx)) - bf1bf2_f[[j]][-(1:N1),1]

	hist(bf12_error, mgp= c(1.7, 0.5, 0), xlab="agp_f - abf_f", ylab="", main="posterior residuals (agp_f - abf_f)", cex.main=0.8)

	#images
	image(xx, xx, matrix(gp1gp2_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="AGP posterior means surface agp_f", cex.main=0.8)

	image(xx, xx, matrix(bf1bf2_f[[j]][-(1:N1),1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="ABF posterior means surface abf_f", cex.main=0.8)

	brk1 <- seq(-max(abs(bf12_error)),max(abs(bf12_error)), length.out= 30)

	image.plot(xx, xx, bf12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.6)
}
```

### 2D Splines-based model

Stancode and standata as a function of the number of knots.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
kn <- c(4,5,6,7,10)  # number of knots

stancode_sp12 <- list()
standata_sp12 <- list()
for(j in 1:length(kn)){  # number of knots

  # Stancode
	stancode_sp12[[j]] <- make_stancode(y ~ t2(x1, x2, k= kn[j]), data = data.frame(x1= c(x1,x_grid[,1]), x2= c(x2,x_grid[,2]), y= c(y, y_grid)))

	# Standata
	standata_sp12[[j]] <- make_standata(y ~ t2(x1, x2, k= kn[j]), data.frame(x1= c(x1,x_grid[,1]), x2= c(x2,x_grid[,2]), y= c(y, y_grid)))

	standata_sp12[[j]]$N1 <- N1
	standata_sp12[[j]]$N_grid <- N_grid
}
```

Load and store the outcomes of the splines-based model as a function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Simulated_data_project/function2/aj_sp12.rData")

sp12_f <- list()
sp12_y <- list()
sp12_elpd <- list()
sp12_time <- list()
for(j in 1:length(kn)){ # number of knots
	sp12_f[[j]] <- summary(aj_sp12[[j]], pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_y[[j]] <- summary(aj_sp12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_elpd[[j]] <- summary(aj_sp12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_time[[j]] <- sum(get_elapsed_time(aj_sp12[[j]])[2])/(aj_sp12[[j]]@sim$iter - aj_sp12[[j]]@sim$warmup)
}
```

Posterior plots of the Splines-based model versus the True function.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots  True function vs SP
for(j in 1:length(kn)){
  par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(f(x1,x2), sp12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="true function f vs SP posterior means sp_f", cex.main=0.9); abline(a=0, b=1, col="gray")
	plot(f(x1,x2), f(x1,x2) - sp12_f[[j]][1:N1,1], xlab="f", ylab="f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (f - sp_f)", cex.main=0.9); abline(h=0, col="gray")

	title(main= paste("True function f vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# predicting
	plot(f(x_grid[,1],x_grid[,2]), sp12_f[[j]][(N1+1):N,1], col=4, cex=0.6, xlab="f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="true function f vs SP posterior means sp_f", cex.main=0.9); abline(a=0, b=1, col="gray")
	plot(f(x_grid[,1],x_grid[,2]), f(x_grid[,1],x_grid[,2]) - sp12_f[[j]][(N1+1):N,1], xlab="f", ylab="f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (f - sp_f)", cex.main=0.9); abline(h=0, col="gray")
	plot(f(x_grid[,1],x_grid[,2]), sp12_elpd[[j]][(N1+1):N,1], col=2, cex=0.8, xlab="f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(sp12_elpd[[j]][(N1+1):N,1]), col="red", lty=2)

	# histogram residuals
	sp12_error <- matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)) - sp12_f[[j]][(N1+1):N,1]

	hist(sp12_error, mgp= c(1.7, 0.5, 0), xlab="f - sp_f", ylab="", main="posterior residuals (f - sp_f)", cex.main=0.9)

	# images
	image(xx, xx, matrix(f(x_grid[,1],x_grid[,2]),length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="True function surface f", cex.main=0.9)

	image(xx, xx, matrix(sp12_f[[j]][(N1+1):N,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="SP posterior means surface sp_f", cex.main=0.9)

	brk1 <- seq(-max(abs(sp12_error)),max(abs(sp12_error)), length.out= 30)

	image.plot(xx, xx, sp12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, main="posterior residuals surface", cex.main=0.9)
	points(x1, x2, pch=1, cex=0.5)
}
```

Posterior plots of the Splines-based model versus the Gaussian process model.

```{r fig.height=8, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots GP vs SP
for(j in 1:length(kn)){
	par(mfcol=c(3,3), mai=c(0.4, 0.5, 0.25, 0.1), omi=c(0,0,0,0.2), oma=c(1, 0, 1.5, 1.5))

	# training
	plot(gp12_f[,1], sp12_f[[j]][1:N1,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs SP posterior means sp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f[,1], gp12_f[,1] - sp12_f[[j]][1:N1,1], xlab="gp_f", ylab="gp_f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - sp_f)", cex.main=0.8); abline(h=0, col="gray")

	title(main= paste("GP vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# predicting
	plot(gp12_f_grid[,1], sp12_f[[j]][(N1+1):N,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", mgp= c(1.7, 0.5, 0), main="GP posterior means gp_f vs SP posterior means sp_f", cex.main=0.8); abline(a=0, b=1, col="gray")
	plot(gp12_f_grid[,1], gp12_f_grid[,1] - sp12_f[[j]][(N1+1):N,1], xlab="gp_f", ylab="gp_f - sp_f", col=4, cex=0.6, mgp= c(1.7, 0.5, 0), main="posterior residuals (gp_f - sp_f)", cex.main=0.8); abline(h=0, col="gray")
	plot(gp12_f_grid[,1], sp12_elpd[[j]][(N1+1):N,1], col=2, cex=0.8, xlab="gp_f", ylab="ELPD", mgp= c(1.7, 0.5, 0)); abline(h=mean(sp12_elpd[[j]][(N1+1):N,1]), col="red", lty=2)

	# histogram residuals
	sp12_error <- matrix(gp12_f_grid[,1],table(x_grid[,2])[1],table(x_grid[,2])[1]) - sp12_f[[j]][(N1+1):N,1]

	hist(sp12_error, mgp= c(1.7, 0.5, 0), xlab="gp_f - sp_f", ylab="", main="posterior residuals (gp_f - sp_f)", cex.main=0.8)

	# images
	image(xx, xx, matrix(gp12_f_grid[,1],length(xx),length(xx)), col=col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="GP posterior means surface gp_f", cex.main=0.8)

	image(xx, xx, matrix(sp12_f[[j]][(N1+1):N,1],length(xx),length(xx)), col= col1, breaks= brk, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), main="SP posterior means surface sp_f", cex.main=0.8)

	brk1 <- seq(-max(abs(sp12_error)),max(abs(sp12_error)), length.out= 30)

	image.plot(xx, xx, sp12_error, col= col3, breaks= brk1, xlab="x1", ylab="x2", mgp= c(1.7, 0.5, 0), legend.width=1.2, main="posterior residuals surface", cex.main=0.8)
	points(x1, x2, pch=1, cex=0.5)
}
```

### Computation time

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:8], c(unlist(bf1bf2_time)[1:8],NA,NA,NA)[1:8], type="b", col=5, ylim=c(0,1.5), xlab="M ; Knots", ylab="seconds per iteration")
lines(M[1:8], c(unlist(bf12_time)[1:7],NA), type="b", col=4)
lines(c(4,6,8,10,16), unlist(sp12_time)[1:5], type="b", col=6)
abline(h= c(gp1gp2_time+0.01, gp12_time), col=3:2)
legend("topright", legend=c("GP  2D","AGP  2D","BF  2D","ABF  2D","SP  2D"), col=2:6, lty=1, cex=0.7)
```

### Application to the Diabetes dataset

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# reference Notebook
# https://rawgit.com/avehtari/modelselection_tutorial/master/diabetes.html

diabetes <- read.csv("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Diabetes/diabetes.csv", header = TRUE)
summary(diabetes)
```

```{r fig.height=3, fig.width=4, message=FALSE, warning=FALSE, paged.print=FALSE}
# removing those observation rows with 0 in any of the variables
for (i in 2:6) {
      diabetes <- diabetes[-which(diabetes[, i] == 0), ]
}
# scale the covariates for easier comparison of coefficient posteriors
for (i in 1:8) {
      diabetes[i] <- as.vector(scale(diabetes[i]))
}

# modify the data column names slightly for easier typing
names(diabetes)[7] <- "dpf"
names(diabetes) <- tolower(names(diabetes))

# bi-plots
p <- ggplot(diabetes, aes(age, glucose))
p + geom_point(aes(colour = outcome), size = 2)

p <- ggplot(diabetes, aes(pregnancies, glucose))
p + geom_point(aes(colour = outcome), size = 2)
```

#### Preparing data for 2-Dimensions

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
### data 2D
set.seed(115)

# indices of test data
ind2 <- sample(1:dim(diabetes)[1], 15)
N2 <- length(ind2)

# indices of training data
ind1 <- setdiff(1:dim(diabetes)[1], ind2)
N1 <- length(ind1)

# grid of predicting data
pred <- pred_grid(coords= c(min(diabetes$pregnancies),max(diabetes$pregnancies)+0.5), y.coords = c(min(diabetes$glucose),max(diabetes$glucose)+0.5), by= 0.5)

# total, training, test and predicting input variables
x1 <- c(diabetes$pregnancies, pred[,1])
x2 <- c(diabetes$glucose, pred[,2])

# response variable
y <- c(diabetes$outcome, rep(0,dim(pred)[1]))

# indices predicting grid
ind_grid <-  setdiff(1:length(x1), 1:(N1+N2))
N_grid <- length(ind_grid)

length(x1)==N1+N2+N_grid
```

### 2D Gaussian process model

Standata y stancode.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Standata
standata_gp12 <- make_standata(y ~ gp(x1,x2, scale= FALSE) - 1 , data = data.frame(x1, x2, y), family= binomial(link = "logit"))

# Stancode
# stancode <- make_stancode(y ~ gp(x1,x2, scale= FALSE) - 1 , data = data.frame(x1, x2, y), family= binomial(link = "logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp12")

standata_gp12$N1 <- N1
standata_gp12$N2 <- N2
standata_gp12$ind1 <- ind1
standata_gp12$ind2 <- ind2
```

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_gp12.rData")

gp12_f <- summary(aj_gp12, pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_y <- summary(aj_gp12, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_finv <- summary(aj_gp12, pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_elpd <- summary(aj_gp12, pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp12_time <- sum(get_elapsed_time(aj_gp12)[2])/(aj_gp12@sim$iter - aj_gp12@sim$warmup)
```

Confusion matrix on training data.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
print(table(as.factor(as.numeric(gp12_finv[ind1,1]>0.5)), y[ind1]))
```

Confusion matrix on test data.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
print(table(as.factor(as.numeric(gp12_finv[ind2,1]>0.5)), y[ind2]))
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
par(mfrow=c(2,2))

hist(as.numeric(gp12_finv[ind1,1]>0.5), main="GP posterior means", xlab="gp_f", cex.main=0.7)
hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

plot(gp12_finv[ind2,1], gp12_elpd[,1], col=4, cex=0.6, xlab="gp_f", ylab="ELPD"); abline(a=0, b=1, col="gray")

n_x <- length(unique(pred[,1]))
n_y <- length(unique(pred[,2]))
z <- matrix(as.vector(gp12_finv[ind_grid,1]), n_x, n_y)
filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), xlab="x1",
plot.axes = {axis(1, round(pred[,1],2)); axis(2, round(pred[,2],2)); points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)])},
plot.title=title(main= "GP posterior mean", xlab="x1", ylab="x2")
)
```

### 2D Basis function model

Preparing data for 2D Basis function approach.

```{r}
D <- 2  # number of dimensions
M <- c(2,4,6,10,20,30,40,60)  # number of basis functions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
			# for (m3 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2) #, m3)
			# }
		}
	}
}

# Standata
standata_bf12 <- list()
for(j in 1:length(M)){  # number of basis functions
  standata_bf12[[j]] <- standata_gp12
	standata_bf12[[j]]$D <- 2
	standata_bf12[[j]]$M <- M[j]
	standata_bf12[[j]]$M_nD <- M[j]^D
	standata_bf12[[j]]$indices <- indices[[j]]
	standata_bf12[[j]]$L <- c(5/2*max(x1),5/2*max(x2))
}
```

Stancode.

(The code of the D-dimensional Basis function aproximation model is missing)

Load and store the outcomes of the different Basis function models as function of the number of basis functions.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_bf12.rData")

bf12_f <- list()
bf12_finv <- list()
bf12_y <- list()
bf12_elpd <- list()
bf12_time <- list()
for(j in 1:length(aj_bf12)){
	bf12_f[[j]] <- summary(aj_bf12[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_finv[[j]] <- summary(aj_bf12[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_y[[j]] <- summary(aj_bf12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_elpd[[j]] <- summary(aj_bf12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf12_time[[j]] <- sum(get_elapsed_time(aj_bf12[[j]])[2])/(aj_bf12[[j]]@sim$iter - aj_bf12[[j]]@sim$warmup)
}
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(10), colorRampPalette(c("white", "red"), bias=10)(10)[2:10])

for(j in 1:length(aj_bf12)){
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

  # scatter  plots of training data
	plot(gp12_finv[ind1,1], bf12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="bf_f", main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.7); abline(a=0, b=1, col="gray")

	plot(gp12_finv[ind1,1], gp12_finv[ind1,1] - bf12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - bf_f", main="posterior residuals (gp_f - bf_f)", cex.main=0.7); abline(h=0, col="gray")

	# histograms
	hist(as.numeric(bf12_finv[[j]][ind1,1]>0.5), main="BF posterior means", xlab="bf_f", cex.main=0.7)
	hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs BF model with M =",M[j]), outer= TRUE)

	# Confussion matrices
	# print(paste("Confussion matrix of BF model with M =",M[j]))
	# print("training data")
	# print(table(as.factor(as.numeric(bf12_finv[[j]][ind1,1]>0.5)), y[ind1]))
	# print("test data")
	# print(table(as.factor(as.numeric(bf12_finv[[j]][ind2,1]>0.5)), y[ind2]))

	# Posterior means
	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(bf12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))
	points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)] )},	plot.title=title(main = paste("bf(x1;x2;M=",M[[j]],sep="",")")) )

	# GP vs BF posterior means
	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(gp12_finv[ind_grid,1]) - as.vector(bf12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, levels= seq(-max(abs(z)),max(abs(z)), length.out= 20), col= col3, plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))},
	plot.title=title(main= paste("gp(x1,x2) - bf(x1;x2;M=",M[[j]],sep="",")")) )
}
```

## 2D Splines-based model

Standata and stancode as a function of the number of knots.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
kn <- c(2,3,4,5,8)  # number of knots

stancode <- list()
standata_sp12 <- list()
for(j in 1:length(kn)){

  # Stancode
	stancode[[j]] <- make_stancode(y | trials(n) ~ t2(x1, x2, k= 2 + kn[j]), data = data.frame(x1= x1, x2= x2, y= y, n= rep(1,N1+N2+N_grid)), family = binomial("logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp12")

	# Standata
	standata_sp12[[j]] <- make_standata(y | trials(n) ~ t2(x1, x2, k= 2 + kn[j]), data = data.frame(x1= x1, x2= x2, y= y, n= rep(1,N1+N2+N_grid)), family= binomial("logit"))

	standata_sp12[[j]]$N1 <- N1
	standata_sp12[[j]]$N2 <- N2
	standata_sp12[[j]]$ind1 <- ind1
	standata_sp12[[j]]$ind2 <- ind2
}

N <- standata_sp12[[1]]$N
```

Load and store de outcomes of the different Splines models as function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_sp12.rData")

sp12_f <- list()
sp12_finv <- list()
sp12_y <- list()
sp12_elpd <- list()
sp12_time <- list()
for(j in 1:length(kn)){ # number of knots
	sp12_f[[j]] <- summary(aj_sp12[[j]], pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_finv[[j]] <- summary(aj_sp12[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_y[[j]] <- summary(aj_sp12[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_elpd[[j]] <- summary(aj_sp12[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	sp12_time[[j]] <- sum(get_elapsed_time(aj_sp12[[j]])[2])/(aj_sp12[[j]]@sim$iter - aj_sp12[[j]]@sim$warmup)
}
```

```{r fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(10), colorRampPalette(c("white", "red"), bias=10)(10)[2:10])

for(j in 1:length(kn)){  # number of knots
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

	plot(gp12_finv[ind1,1], sp12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="sp_f", main="GP posterior means gp_f vs SP posterior means sp_f", cex.main= 0.7); abline(a=0, b=1, col="gray")

	plot(gp12_finv[ind1,1], gp12_finv[ind1,1] - sp12_finv[[j]][ind1,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - sp_f", main="posterior residuals (gp_f - sp_f)", cex.main= 0.7); abline(h=0, col="gray")

	hist(as.numeric(sp12_finv[[j]][ind1,1]>0.5), main="SP posterior means", xlab="sp_f", cex.main=0.7)
	hist(y[ind1], main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs SP model with knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3]), outer= TRUE)

	# print(paste("Confussion matrix of the Splines model with Knots =",kn[j]))
	# print("training data")
	# print(table(as.factor(as.numeric(sp12_y[[j]][ind1,1]>0.5)), y[ind1]))
	# print("test data")
	# print(table(as.factor(as.numeric(sp12_y[[j]][ind2,1]>0.5)), y[ind2]))

	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(sp12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, zlim= range(z, finite= TRUE), levels= pretty(range(z, finite = TRUE), n = 10), col= cm.colors(length(pretty(range(z, finite = TRUE), n = 10)) - 1, alpha=0.5), plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))
	points(x=diabetes$pregnancies, y=diabetes$glucose, pch=16, cex=0.7, col=cm.colors(2)[cut(y, breaks=2)] )},	plot.title=title(main = paste("sp(x1;x2;Knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3],")")) )

	n_x <- length(unique(pred[,1]))
	n_y <- length(unique(pred[,2]))
	z <- matrix(as.vector(gp12_finv[ind_grid,1]) - as.vector(sp12_finv[[j]][ind_grid,1]), n_x, n_y)
	filled.contour(x = unique(pred[,1]), y = unique(pred[,2]), z= z, asp= 1, levels= seq(-max(abs(z)),max(abs(z)), length.out= 20), col= col3, plot.axes = {axis(1, round(pred[,1],2))
	axis(2, round(pred[,2],2))},
	plot.title=title(main= paste("gp(x1,x2) - sp(x1;x2;Knots =",standata_sp12[[j]]$knots_1[1],standata_sp12[[j]]$knots_1[2],standata_sp12[[j]]$knots_1[3],")")) )
}
```

### Computation time

```{r echo=TRUE, fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:8], c(unlist(bf12_time)[1:5],NA,NA,NA)[1:8], type="b", col=4, ylim=c(0,gp12_time+1), xlab="M ; Knots", ylab="seconds per iteration")
lines(kn[1:5], unlist(sp12_time)[1:5], type="b", col=6)
abline(h= gp12_time, col=2)
legend("bottomright", legend=c("GP  2D","BF  2D","SP  2D"), col=c(2,4,6), lty=1, cex=0.7)
```

#### Preparing data for 3-Dimensions

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
### data 3D
set.seed(115)
ind2 <- sample(1:dim(diabetes)[1], 15)
N2 <- length(ind2)

ind1 <- setdiff(1:dim(diabetes)[1], ind2)
N1 <- length(ind1)

x1 <- diabetes$pregnancies
x2 <- diabetes$glucose
x3 <- diabetes$age
y <- diabetes$outcome

length(x1)==N1+N2
```

### 3D Gaussian process model

Standata y stancode.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Standata
standata_gp123 <- make_standata(y ~ gp(x1,x2,x3, scale= FALSE) - 1 , data = data.frame(x1, x2, x3, y), family= binomial(link = "logit"))

# Stan code
# stancode <- make_stancode(y ~ gp(x1,x2,x3, scale= FALSE) - 1 , data = data.frame(x1, x2, x3, y), family= binomial(link = "logit")) #, save_model= "C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/stancode_gp123")

standata_gp123$N1 <- N1
standata_gp123$N2 <- N2
standata_gp123$ind1 <- ind1
standata_gp123$ind2 <- ind2
```

Load and store the outcomes of the GP model.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_gp123.rData")

gp123_f <- summary(aj_gp123, pars = c("mu"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_y <- summary(aj_gp123, pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_finv <- summary(aj_gp123, pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_elpd <- summary(aj_gp123, pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
gp123_time <- sum(get_elapsed_time(aj_gp123)[2])/(aj_gp123@sim$iter - aj_gp123@sim$warmup)
```

Confussion matrix of the GP model for training data.

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
table(as.factor(as.numeric(gp123_y[,1]>0.5)), y)
```

Confussion matrix of the GP model for test data.

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
table(as.factor(as.numeric(gp123_y[ind2,1]>0.5)), y[ind2])
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
par(mfrow=c(2,2))

hist(as.numeric(gp123_y[,1]>0.5), main="GP posterior means", xlab="gp_y", cex.main=0.7)
hist(y, main="observed data", xlab="y", cex.main=0.7)

plot(gp123_y[ind2,1], gp123_elpd[,1], col=4, cex=0.6, xlab="gp_y", ylab="ELPD"); abline(a=0, b=1, col="gray")
```

### 3D Basis function model

Standata.

```{r}
D <- 3  # number of dimensions
M <- c(2,4,6,10,20,30,40,60)  # number of basis functions

indices <- list()
for(j in 1:length(M)){
	indices[[j]] <- matrix(NA, M[j]^D, D)
	mm=0;
	for (m1 in 1:M[j]){
		for (m2 in 1:M[j]){
			for (m3 in 1:M[j]){
				mm = mm+1
				indices[[j]][mm,] = c(m1, m2, m3)
			}
		}
	}
}

standata_bf123 <- list()
for(j in 1:length(M)){
  standata_bf123[[j]] <- standata_gp123
	standata_bf123[[j]]$D <- D
	standata_bf123[[j]]$M <- M[j]
	standata_bf123[[j]]$M_nD <- M[j]^D
	standata_bf123[[j]]$indices <- indices[[j]]
	standata_bf123[[j]]$L <- c(5/2*max(x1),5/2*max(x2),5/2*max(x3))
}
```

Stancode.

(The code of the D-dimensional Basis fucntion aproximation model is missing2)

Load and store the outcomes of the Splines-based model model as function of the number of knots.

```{r}
load("C:/GABRIEL_20180206/GIFLE/Proyecto-GAMs/Basisfunctions/BF-2D/Diabetes_project/aj_bf123.rData")

bf123_f <- list()
bf123_finv <- list()
bf123_y <- list()
bf123_elpd <- list()
bf123_time <- list()
for(j in 1:length(aj_bf123)){
	bf123_f[[j]] <- summary(aj_bf123[[j]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_finv[[j]] <- summary(aj_bf123[[j]], pars = c("f_invlogit"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_y[[j]] <- summary(aj_bf123[[j]], pars = c("y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_elpd[[j]] <- summary(aj_bf123[[j]], pars = c("log_y_predict"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
	bf123_time[[j]] <- sum(get_elapsed_time(aj_bf123[[j]])[2])/(aj_bf123[[j]]@sim$iter - aj_bf123[[j]]@sim$warmup)
}
```

```{r echo=TRUE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
# Plots
for(j in 1:length(aj_bf123)){
	par(mfrow=c(2,2), oma=c(1, 0, 1.5, 1.5))

	plot(gp123_finv[,1], bf123_finv[[j]][,1], col=4, cex=0.6, xlab="gp_f", ylab="bf_f", main="GP posterior means gp_f vs BF posterior means bf_f", cex.main=0.7); abline(a=0, b=1, col="gray")

	plot(gp123_finv[,1], gp123_finv[,1] - bf123_finv[[j]][,1], col=4, cex=0.6, xlab="gp_f", ylab="gp_f - bf_f", main="posterior residuals (gp_f - bf_f)", cex.main=0.7); abline(h=0, col="gray")

	hist(as.numeric(bf123_y[[j]][,1]>0.5), main="BF posterior means", xlab="bf_y", cex.main=0.7)
	hist(y, main="observed data", xlab="y", cex.main=0.7)

	title(main= paste("GP vs BF model with M =",M[j]), outer= TRUE)

	print(paste("Confussion matrix of the Basis function model with M =",M[j]))
	print("training data")
	print(table(as.factor(as.numeric(bf123_y[[j]][,1]>0.5)), y))
  print("test data")
	print(table(as.factor(as.numeric(bf123_y[[j]][ind2,1]>0.5)), y[ind2]))
}
```

### Computational time

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(M[1:6], c(unlist(bf123_time)[1:5],NA,NA,NA,NA)[1:6], type="b", col=4, ylim=c(0,gp123_time+0.5), ylab="seconds per iteration", xlab="M")
abline(h= c(gp123_time+0.01), col=5)
legend("topright", legend=c("GP  3D","BF  3D"), col=5:4, lty=1, cex=0.7)
```