### 2020-09-10 Paper/StanCon/poster/HSGP_StanCon_poster_landscape.tex

\documentclass[landscape,a1,final]{a0poster} % a0poster class sets the paper size
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[SCI]{aaltologo} % RGB, Coated, Uncoated
%\usepackage{times}
%\usepackage{lastpage}
\usepackage[center]{titlesec} % For changing the font on chapters, sections, etc.
\usepackage{amsmath,amsfonts,amssymb,amsbsy}
\usepackage{soul}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{amsfonts}
\usepackage{amsmath}
%\usepackage{sectsty}
\usepackage{subcaption}
\usepackage{url}
\usepackage{bm}
\usepackage[colorlinks,linkcolor=black,citecolor=black, urlcolor=blue,filecolor=blue]{hyperref}
\usepackage[left=0.5cm,right=0.5cm,bottom=1cm,top=1.5cm]{geometry} 
\usepackage{changepage}
\usepackage{amsthm}
\usepackage{tcolorbox}
\usepackage{pbox}
\usepackage{multicol}
\usepackage{nicefrac}
\usepackage{setspace}
\usepackage{bibentry}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{,}{,}

\newtheorem*{remark}{Remark}

%%%%%%titleformat
\newcommand{\cajados}[1]{\fcolorbox{aaltoLightGreen!30}{aaltoLightGreen!30}
{\parbox[c][8mm][c]{0.95\textwidth}
{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesection. #1}}}

\titleformat{\section}[block]
{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\cajados}

%\titleformat{\section}{\centering\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{aaltoPurple}{\thesection}}{1em}{}

%\newcommand{\caja_sub}[1]{\fcolorbox{aaltoLightGreen!10}{aaltoLightGreen!10}
%{\parbox[c][8mm][c]{0.95\textwidth}
%{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesubsection. #1}}}
%
%\titleformat{\subsection}[block]
%{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\caja_sub}

\titleformat{\subsection}{\normalsize\sc\bfseries\sffamily\color{aaltoPurple}}{\textcolor{black}{\thesubsection}}{1em}{}

\renewcommand{\emph}[1]{\sffamily{\textcolor{aaltoPurple}{#1}}}

% Beamer-style math font
%\usepackage{fouriernc}
\usepackage{sfmath}

% redefine the bullet symbol 
\renewcommand\labelitemi{\textcolor{aaltoPurple}{$\blacktriangleright$}\,}

% reduce the space between bullets
\let\tempone\itemize
\let\temptwo\enditemize
\renewenvironment{itemize}{\tempone\addtolength{\itemsep}{-0.3\baselineskip}}{\temptwo}

%%%%%%%%%%%%%%%%%%% for tikz figures %%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{ifthen}
\usepackage{tikz}
\usepackage{pgfplots}  
\usetikzlibrary{positioning} 
% \usepgfplotslibrary{external}
%\tikzexternalize

%\usetikzlibrary{matrix}
%\usetikzlibrary{calc}
\newlength{\figurewidth}
\newlength{\figureheight}

\def\figpdfdir{fig/} % directory for pdf-figures
\def\figtikzdir{tikz/} % directory for tikz-figures

% this is replacement for the \input command used in the figure-environment which
% takes into account whether pdf is forced
\newcommand{\minput}[2][]{
\ifthenelse{\equal{#1}{pdf}}
	{ \includegraphics{\figpdfdir #2} }
	{ \tikzset{external/remake next} \tikzsetnextfilename{#2} \input{\figtikzdir #2} }
}

% for externalization
% \usetikzlibrary{external}
% \tikzexternalize[prefix=\figpdfdir]
% \tikzset{external/system call={lualatex
% 	\tikzexternalcheckshellescape -halt-on-error -interaction=batchmode
% 	-jobname "\image" "\texsource"}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\vc}[1] { #1 } %
\newcommand{\vs}[1] { #1 }
\newcommand{\tp}{\mathsf{T}}
\newcommand{\ti}[1] { \tilde{#1} }
\newcommand{\mc}[1] { \mathcal{#1} } 
\newcommand{\tx}[1] { \text{#1} } 
\newcommand{\given} { \,|\, }
\newcommand{\data} {D}
\newcommand{\subjto}{ \, \tx{s.t.} }
\newcommand{\tr}{^{\text{T}} }
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}

%
\newcommand{\pr}[1]{ \Pr {\left[#1\right]} }
\newcommand{\partiald}[2]{\frac{\partial {#1}}{\partial {#2}}}
\newcommand{\mean}[2][] { \mathrm{E}_{#1} {\left[#2\right]} }
\newcommand{\Var}[1] { \mathrm{Var} {\left[#1\right]} }
\newcommand{\var}[1] { \mathrm{var} {\left(#1\right)} }
\newcommand{\Cov}[1] { \mathrm{Cov} {\left[#1\right]} }
%\newcommand{\cov}[1] { \mathrm{cov} {\left({#1}\right)}}
\newcommand{\Tr}[1] { \mathrm{Tr} {\left[#1\right]} }
\newcommand{\KL}[2] { \mathrm{KL} {\left(#1 \, \| \, #2\right)} }
\newcommand{\x}{\mathbf{x}}
\newcommand{\cx}{\mathbf{c}}
\newcommand{\f}{\mathbf{f}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\F}{\mathbf{F}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\id}{\text{d}}
%
%\newcommand{\Normal}[1] { \mathrm{N} {\left(#1\right)}  } 
\newcommand{\Student}[2][] { t_{#1} {\left(#2\right)} }
\newcommand{\halfStudent}[2][] { t_{#1}^+ {\left(#2\right)} }
\newcommand{\SInvchi}[1] { \mathrm{Scale\text{-}Inv\text{-}}\chi^2 {\left(#1\right)} }
\newcommand{\InvGamma}[1] { \mathrm{Inv\text{-}Gamma} {\left(#1\right)} }
\newcommand{\Cauchy}[1] { \mathrm{C} {\left(#1\right)} }
\newcommand{\halfCauchy}[1] { \mathrm{C}^+ {\left(#1\right)} }
\newcommand{\Beta}[1] {\mathrm{Beta}{\left(#1\right)} }
\newcommand{\Ber}[1] {\mathrm{Ber}{\left(#1\right)} }

%
\newcommand{\NormalNP} { \mathrm{N} } 
\newcommand{\StudentNP}[1][] { t_{#1} }
\newcommand{\halfStudentNP}[1][] { t_{#1}^+ }
\newcommand{\SInvchiNP} { \mathrm{Scale\text{-}Inv\text{-}}\chi^2  }
\newcommand{\InvGammaNP} { \mathrm{Inv\text{-}Gamma} }
\newcommand{\CauchyNP} { \mathrm{C}  }
\newcommand{\halfCauchyNP} { \mathrm{C}^+ }
\newcommand{\BetaNP} {\mathrm{Beta} }
\newcommand{\BerNP} {\mathrm{Ber} }

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator{\GP}{GP}
%\DeclareMathOperator{\Normal}{Normal}
\DeclareMathOperator{\cov}{cov}

% --- Acronyms
\usepackage{xspace}
\newcommand{\acr}[1]{\textsc{#1}\xspace}
\newcommand{\gp}{\acr{gp}}
\newcommand{\ep}{\acr{ep}}
\newcommand{\gps}{\acr{gp}{\!\,}s }
\newcommand{\abc}{\acr{abc}}
\newcommand{\bo}{\acr{bo}}
% optimization

\definecolor{navyblue}{rgb}{0,0,0.55}
\definecolor{lightblue}{rgb}{0.91,0.91,0.97}

\newcommand{\sectionspace}{0em} % Free space before each section inside a minipage
\newcommand{\figurespace}{0.5em} % Free space around figures inside a minipage (where floats are not allowed)

% Reduce the vertical space taken by the equations
\setlength{\abovedisplayskip}{10pt}
\setlength{\belowdisplayskip}{10pt}
%\setlength{\abovedisplayshortskip}{10pt}
%\setlength{\belowdisplayshortskip}{10pt}

%% Font sizes for a0poster are
%\tiny
%\scriptsize
%\footnotesize
%\small
%\normalsize
%\large
%\Large
%\LARGE
%\huge
%\Huge
%\veryHuge
%\VeryHuge
%\VERYHuge

%% Official colors from aaltologo-package (visual-identity guideline)
% aaltoBlack
% aaltoGray
% aaltoGrayScale (for b&w prints)
% aaltoYellow
% aaltoOrange
% aaltoRed
% aaltoFuchsia
% aaltoPurple
% aaltoBlue
% aaltoTurquoise
% aaltoGreen
% aaltoLightGreen
% You can use \textcolor{<aaltocolor>}{<your text>) to change the colors of text, or

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\thispagestyle{empty} % Removes the page number

%------------ minipage for the logo & title ------------%
\begin{minipage}[t]{1\linewidth} 
\vspace{0pt} % A trick to align the parallel minipages on top

\centering

\vspace{-0.015\linewidth}

%-- minipage for the logo at left side --%
\begin{minipage}[t]{0.14\linewidth} % logo
\vspace{0.5cm}

%\begin{adjustwidth}{-1cm}{-0cm}
\includegraphics[scale=0.35]{\figpdfdir StanCon_logo.png}
%\end{adjustwidth}

\end{minipage} % no empty line before the next begin
%-- minipage for title and authors --%
\begin{minipage}[t]{0.50\linewidth}
\vspace{1cm}

\centering
{\renewcommand{\baselinestretch}{0.6} 
\textcolor{aaltoGreen}{\huge{\bfseries{\textsf{Practical implementation of Hilbert space reduced-rank approximate Gaussian processes} }}}
\par}

\vspace{5mm}
\hspace{-1cm}
\normalsize{\textsf{\bfseries{Gabriel Riutort-Mayol$^{1}$ \;\; Paul-Christian Bürkner$^2$ \;\; Michael Riis Andersen$^{3}$ \;\; Arno Solin$^{2}$ \;\; Aki Vehtari$^{2}$}}}

\end{minipage}
%-- minipage for the logo at rigth side --%
\begin{minipage}[t]{0.22\linewidth}

\vspace{-2mm}
\textcolor{aaltoGray}{\textsf{\bfseries{ \footnotesize
\begin{itemize}
	\setlength{\itemsep}{0mm}
	\item[] $^1$ Department of Cartographic Engineering, Geodesy, and Photogrammetry, Universitat Polit\`ecnica de Val\`encia, Spain
\item[] $^2$ Department of Computer Science, Aalto University, Finland
\item[] $^3$ Department of Applied Mathematics and Computer Science, Technical University of Denmark, Denmark
\vspace{2mm}
\item[] {\small Contact: gabriuma@gmail.com,\, aki.vehtari@aalto.fi}
\end{itemize}
}}}

\end{minipage}
%-- minipage for the logo at rigth side --%
\hspace{15mm}
\begin{minipage}[t]{0.09\linewidth}
\vspace{10mm}

%\begin{adjustwidth}{-1cm}{-0cm}
\includegraphics[scale=0.8]{\figpdfdir aaltologo}
\raisebox{-5mm}{\includegraphics[scale=0.7]{\figpdfdir marca_UPV_secundaria_negro300.png}}
%\end{adjustwidth}
\end{minipage}

\end{minipage}

% Default font for text
\sffamily
\normalsize % Text size
%\small % Text size
%\footnotesize % Text size

\vspace{1cm}
This poster is based on the pre-print publication at {\url{https://arxiv.org/abs/2004.11408}}

% Space according to the visual identity guidelines...
%\vspace{0.02\linewidth}

% Centering helps in placement
\centering

%------------ minipage for the main text ------------%
\begin{minipage}{1\linewidth} 
%------------ minipage for first column ------------%
%\hspace{0.001\textwidth}
\centering
\hspace{3pt}
\begin{minipage}[t]{0.30\linewidth}
\vspace{0pt}
\setlength{\parindent}{10mm} % Paragraph indent

\vspace{0.8cm}
\section{Introduction}
\vspace{-0.2cm}

\begin{itemize}
	\setlength{\itemsep}{3mm}
%	\item Gaussian processes (GPs) are powerful non-parametric probabilistic models for stochastic functions.
	\item When Gaussian processes are implemented using covariance matrices, the computation cost of evaluating the log posterior density is\, $O(n^3)$\, for\, $n$\, observations, which makes them computationally intractable with moderate or large datasets.
	\item We focus on a novel approach for reduced-rank approximate GPs, based on a basis function approximation via Laplace eigenfunctions
for stationary covariance functions \href{https://arxiv.org/abs/1401.5508}{\citep{solin2018hilbert}}. We call this reduced-rank GP as {\color{navyblue} Hilbert space approximate Gaussian process} (HSGP).
	\item This HSGP is simple in complexity due to its linear structure\; $\to$\; Improves sampling efficiency and makes inference considerably faster.

\end{itemize}

\vspace{-7mm}
\centering  %%blue!20; ligthblue
\hspace{-2mm}
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} Our contribution:} Performance analysis for practical implementation of this HSGP, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=1\textwidth]
	\begin{itemize}\setlength\itemsep{2mm}
	\item {\color{navyblue} Discover the relationships} among the key factors of the method such as the number of basis functions, domain of the prediction space, and wiggliness of the function to be learned.
	\item {\color{navyblue} Practical recommendations} for choosing the values of these key factors based on these relationships. %to improve performance and save computation time
	\item {\color{navyblue} Diagnosis} for checking that the number of basis functions and the domain of the prediction space are adequate given the data.
	\end{itemize}
\end{tcolorbox}

\vspace{-0.7cm}
\section{Gaussian process as a prior for latent functions}

\begin{itemize}\setlength\itemsep{3mm}
\item Function values\, $\bm{f}=\left\lbrace  f(\bm{x}_1), f(\bm{x}_2), \hdots, f(\bm{x}_n) \right\rbrace$,\, following a GP model has the defining property:
%\vspace{-4mm}
\begin{align*}
p(\bm{f}) &= \mathcal{N}(\bm{f} \mid \bm{\mu}, K),\\[1mm]
\mu_i &= m(\bm{x}_i),\\
\hspace{27mm} K_{i,j} &=\cov( f(\bm{x}_i),f(\bm{x}_j))=k(\bm{x}_i,\bm{x}_j).
\end{align*}

\item The covariance function encodes our prior assumptions about the variation of the function and defines the {\color{navyblue} correlation structure} of function values \citep{rasmussen2006gaussian}.

\item The main parameter of covariance functions from the Matern class is the {\color{navyblue} length-scale}\, $\ell > 0$, which controls the {\color{navyblue} smoothness/wigglyness} of the function samples.

%\item Function samples (from a squared exponential covariance function) for different lengthscale\, $\ell$:\: 
\end{itemize}

\begin{flushleft}
\hspace{40mm} $\ell=0.1$ \hspace{48mm} $\ell=0.3$ \hspace{47mm} $\ell=0.6$
\end{flushleft}
\vspace{-3mm}
\includegraphics[scale=0.45, trim = 0mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_1.pdf}
\includegraphics[scale=0.45, trim = 34mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_2.pdf}
\includegraphics[scale=0.45, trim = 34mm 0mm 7mm 9mm, clip]{\figpdfdir func_samples_3.pdf}

\end{minipage}% end minipage first column
\hspace{0.015\linewidth} 
%------------ minipage for second column ------------%
\begin{minipage}[t]{0.31\textwidth}
\vspace{0pt}
\setlength{\parindent}{10mm}

\vspace{0.3cm}
\section{Hilbert space approximate Gaussian process (HSGP)}
\vspace{-0.2cm}

\begin{itemize}\setlength\itemsep{3mm}

\item The covariance function is represented as a {\color{navyblue} infinite series expansion} of eigenfunctions\, $\phi_j(x)$\, scaled by the spectral density\, $S_{\theta}(\sqrt{\lambda_j})$\, {\color{darkgray} \citep{solin2018hilbert}}, and {\color{navyblue} approximated by the first $m$ terms}:
%
\begin{align*}
k(x,x') = \sum_{j=1}^m S(\sqrt{\lambda_j}) \phi_j(x) \phi_j(x'), \hspace{10mm} x,x' \in [-L,L]
\end{align*}

\item where $S(\cdot)$ is the {\color{navyblue}spectral density function} of the covariance function $k(\cdot,\cdot)$.

%\begin{remark}
%Any stationary covariance function\, $k:\bm{x} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, and its {\color{navyblue} spectral density}\, $S:\bm{w} \in {\rm I\!R}^{D} \to {\rm I\!R}$\, are Fourier duals, where\, $\bm{w}$\, is a frequency. So, stationary covariance functions can be represented in terms of their spectral density functions.
%\end{remark}

\item We choose the {\color{navyblue} eigenvalues}\, $\lambda_j$\, and {\color{navyblue} eigenfunctions}\, $\phi_j(x)$\, of the Laplacian $-\nabla^2$:
%
\begin{align*}
\left. \begin{array}{cc}
 -\nabla^2 \phi_j(x) = \lambda_j \phi_j(x), \hspace{0.8cm}  x \in [-L,L] \\
\hspace{6mm} \phi_j(x) = 0, \hspace{2cm} x \notin [-L,L] 
\end{array} \right\} \Rightarrow \begin{array}{ll}
 \lambda_j=\left(\frac{j\pi}{2L}\right)^2 \\
\phi_j(x)=\sqrt{\frac{1}{L}} \text{sin}\left(\sqrt{\lambda_j}(x+L)\right) 
\end{array}
\end{align*}

\item $L > 0$\, represents the {\color{navyblue} boundaries} where this representation is valid.

%\item $S(\sqrt{\lambda_j})$\, is monotonically decreasing with\, $j$, and decays rapidly to zero, and this rate of decay depends on the lenghtscale of\, $k$\, (smoothness of the function to be learned) \; $\rightarrow$ \; This points to {\color{navyblue} a finite number}\, $m$\, {\color{navyblue} of terms for an accurate approximation}, as long as\, $x_i$\, are not too close to the boundaries\, $-L$\, and\, $L$\, of\, $\Omega$.

\item This representation of the covariance function\, $k(x,x')$\, leads to the {\color{navyblue} linear representation} of the function\, $f$:
%
\begin{align*}
f(x) \approx \sum_{j}^m \phi_j(x) \beta_j, \hspace{10mm} \beta_j \sim \mathcal{N}(0,S(\sqrt{\lambda_j})).
\end{align*}
\end{itemize}

\vspace{-0.5cm}
\begin{tcolorbox}[colframe=blue!20, colback=white, title=\small \color{black!80} Properties of the HSGP model, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt]
\begin{itemize}\setlength\itemsep{1mm}
\item[+] The basis functions\, $\phi_j(x)$\, do not depend on the hyperparameters, so they just need to be computed once with cost\, $O(m^2n)$.

\item[+]  All dependencies on the hyperparameters is through the variance of the weights\, $\beta_j$. The computational cost of evaluating the log posterior density is\, $O(mn+m)$, which contrasts with\, $O(n^3)$\, of exact GPs.

%\item[+] $f(x)$ is naturally in the {\color{navyblue} non-centered parameterization form} with independent prior distribution on $\beta_j$ which makes posterior inference easier. \citep[see, e.g., ][]{Betancourt+Girolami:2019}

\item[+] The parameter posterior\, $p(\bm{\beta}|\bm{y})$\, is\, $m$-dimensional ($m\ll n$), which makes posterior inference easier.

\item[+] Reduced memory requirements of autodiff methods used in Stan software.

%\item[+] Trade-off between accuracy and computation can also be carried out.

\item[-] The number of multivariate basis functions,\, $m^{\ast} = \prod_{d=1}^{D} m_d$,\, grows exponentially with respect to\, $D$. In practice,\, $D>3$\, starts to be too computationally demanding, even for smooth functions.

\item[+] In high dimensional cases, HSGP may still be used for low-dimensional components in an additive modeling scheme as complexity is linear with the number of additive components.

\end{itemize}
\end{tcolorbox}

\vspace{-0.5cm}
\section{Performance analysis of HSGP}
\vspace{-0.2cm}

\begin{itemize}\setlength\itemsep{2mm}
\item Determine the minimum\, $m$\, ({\color{navyblue} basis functions}) and valid\, $L$\, ({\color{navyblue} boundary}) to obtain accurate approximations, as a function of\, $\ell$\, ({\color{navyblue} lengthscale}).

\end{itemize}

%% bibliography
%\rule{\textwidth}{1mm}
%{\centering\textbf{References}}
%\renewcommand{\bibsection}{}
%{\footnotesize 
\bibliographystyle{plainnat} %unsrt
\nobibliography{references}
%}

\end{minipage}% end minipage second column
\hspace{0.015\linewidth} 
%------------ minipage for third column ------------%
\begin{minipage}[t]{0.345\textwidth}
\vspace{0pt}
\setlength{\parindent}{10mm}

\vspace{-0.8cm}
\hspace{-1cm}
%\subsection{Model of relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}
\textsc{\textbf{4.1 Model of relationships among $m$, $L$ and $\ell$ in terms of an accurate approximation}}

\vspace{4mm}
\hspace{-1.3cm}
\begin{minipage}{0.32\textwidth}
\hspace{0.22\textwidth} \scriptsize Squared exponential kernel\\
% \hspace{0.3\textwidth}  \scriptsize Mattern($\nu$=3/2) kernel\\
\includegraphics[width=0.98\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig6_lscale_vs_J_vs_c_zoomin.pdf} \;\;
%\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig6_lscale_vs_J_vs_c_zoomin_Matern.pdf}

\end{minipage}
\begin{minipage}{0.67\textwidth}

\vspace{-1.5cm}
\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} This model of relationships says...}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.975\textwidth]
\begin{itemize}\setlength\itemsep{1mm}
\item For a given\, $m$\, and\, $c$,\, the minimum\, $\ell$\, that can be accurately approximated.
\item For a given\, $\ell$\, and\, $c$,\, the minimum\, $m$\, required for an accurate approximation.
\item For a given\, $\ell$\, and\, $m$,\, the highest\, $c$\, allowed for an accurate approximation.
\item The minimum valid\, $c$\, for a given\, $\ell$.
\end{itemize}
\end{tcolorbox}
\end{minipage}

\vspace{-3mm}
{\centering
\begin{tcolorbox}[colframe=blue!20, colback=white, title={\small \color{black!80} This model of relationships serves as a\, \textsc{\textbf{\color{black} \normalsize diagnosis tool}}\, of the approximation:}, colbacktitle=lightblue, coltitle=black, boxrule=0.5pt, width=0.98\textwidth]
\begin{remark}
As the lengthscale is the parameter that ultimately characterizes the non-linearity of the posterior functions, we assume we can look at the {\color{navyblue} estimate lengthscale}\, $\hat{\ell}$\, to diagnose the approximation.
\end{remark}

\begin{itemize}\setlength\itemsep{1mm}
\item Given $m$ and $c$

\item If\, $\hat{\ell}$\, {\color{navyblue} is below} the minimum\, $\ell$\, that can be accurately approximated (for those given\, $m$\, and\, $c$), then the approximation {\color{navyblue} is not sufficiently accurate}, and\, $m$\, must be increased or\, $c$\, decreased.

\item If\, $\hat{\ell}$\, {\color{navyblue} is above} the minimum\, $\ell$,\, then the approximation should be {\color{navyblue} sufficiently accurate}.

\item This can be done iteratively.

%\item $c$\, can not be decreased as much as desired because it is restricted to\, $\ell$.
\end{itemize}
\end{tcolorbox}
}

%%%%%%titleformat
\newcommand{\pepito}[1]{\fcolorbox{aaltoLightGreen!30}{aaltoLightGreen!30}
{\parbox[c][16mm][c]{0.95\textwidth}
{\centering\normalsize\sc\bfseries\sffamily\color{black} \thesection. #1}}}

\titleformat{\section}[block]
{\normalfont\filcenter\color{aaltoGray}}{}{1em}{\pepito}
%%%%%%

\begin{minipage}{0.35\textwidth}

\vspace{-2cm}
\section{Univariate synthetic case study}
\vspace{-0.5cm}
\begin{itemize}\setlength\itemsep{3mm}
\small
%\item \textit{\scriptsize Samples draws from}:%\\ $x \in [-1,1] \subset {\rm I\!R}$,
%\\ GP Mattern($\nu$=3/2, $\ell$=0.15, $\alpha$=1)\\ + $\bm{\epsilon} \sim \mathcal{N}(0, 0.2^2)$

%\item \textit{\scriptsize Observation model}: \pbox{6cm}{$\bm{y} = \bm{f} + \bm{\epsilon}$ \\
%$\bm{\epsilon} \sim \mathcal{N}(\bm{0}, \sigma^2 \bm{I})$}

\item \pbox{8cm}{\textit{\footnotesize Exact GP prior for $\bm{f}$}:\\ $f(x) \sim \GP(0, k_{\nicefrac{3}{2}}(x, x'))$}

\item \pbox{11cm}{\textit{\footnotesize HSGP prior for $\bm{f}$}:\\ $f(x) \approx \sum_{j=1}^m \phi_j(x) \beta_j$,\\ 
$\beta_j \sim \mathcal{N}(0,S_{\nicefrac{3}{2}}(\sqrt{\lambda_j}))$}
%\\ $m=60$, $c=1.5$}
\end{itemize}

\end{minipage}
\begin{minipage}{0.62\textwidth}
\centering
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir ch5_fig10_Posteriors_exI_vposter.pdf}
\end{minipage}

\vspace{-0.8cm}
\hspace{-10mm} \textsc{\textbf{\normalsize Diagnosis}}: {\small(iterative procedure)}\\[-18pt]

\hspace{-1cm}
\begin{minipage}[t]{0.48\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 1}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=20$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.09$}

\vspace{0.3cm}
%{\hspace{2.2cm} \scriptsize Error}\\[-3mm]
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig2_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_1_1_vposter.pdf}

\end{minipage}

\end{minipage}
\hspace{0.5cm}
\begin{minipage}[t]{0.48\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 2}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=30$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.12$}

\vspace{0.3cm}
\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig3_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig6_2_1_vposter.pdf}

\end{minipage}

\end{minipage}

\vspace{5mm}
\hspace{-1cm}
\begin{minipage}[t]{0.48\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 3}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=40$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.14$}

\vspace{0.3cm}
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig4_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  \tiny Mattern($\nu$=3/2) kernel\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_3_1_vposter.pdf}

\end{minipage}

\end{minipage}
\hspace{0.5cm}
\begin{minipage}[t]{0.48\textwidth}

\begin{minipage}[t]{0.45\textwidth}
{\large 4}\hspace{0.1cm} \pbox{2cm}{\footnotesize $m=60$\\ $c=1.5$} $\to$ {\footnotesize$\hat{\ell}=0.14$}

\vspace{0.3cm}
\includegraphics[width=1\textwidth, trim = 0mm 0mm 5mm 10mm, clip]{\figpdfdir caseI_fig5_vposter.pdf}

\end{minipage}
\begin{minipage}[t]{0.50\textwidth}
\vspace{-0.5cm}
\hspace{0.25\textwidth}  {\tiny Mattern($\nu$=3/2) kernel}\\
\includegraphics[width=1\textwidth, trim = 0mm 0mm 4mm 10mm, clip]{\figpdfdir caseI_fig6_4_1_vposter.pdf}

\end{minipage}

\end{minipage}

\end{minipage} % End minipage third column
\hspace{0.02\linewidth} 
\end{minipage} % end minipage main text

\end{document}