### 2022-02-16 Paper/Case-study_Birthday-data/Rnotebook_birthday.Rmd

---
title: 'Case study: Birthday data'
author: "G. Riutort-Mayol, M.R. Andersen, P.C. Burkner, A. Solin and A. Vehtari"
date: "`r format(Sys.Date())`."
output:
  html_document:
    theme: readable
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_download: yes
---

<style type="text/css">
body, td{ font-size: 16px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
</style>

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(ggplot2)
library(loo)
library(latex2exp)
```

## Data

```{r message=FALSE, warn=FALSE}
data <- read.csv(file="../births_usa_1969.csv")
str(data)
```

```{r message=TRUE, warn=TRUE, fig.height=5, fig.width=8}
par(mai=c(1.02,1.2,0.3,0.42))

ind <- data$id
births <- data$births
labels_at = aggregate(data, by=list(data$year), FUN=min)$id

plot(ind, births[ind], type="p", pch=20, bg=grey(0.4), cex=0.6, col=grey(0.5), xlab="", ylab="", lwd=1, mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.2, cex.lab=1.2, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")
axis(1, at = labels_at, c("1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988"), lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.2, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.2, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Year", mgp=c(2.2, 1, 0), cex.lab=1.2, las=1)
title(ylab="Births", mgp=c(3.7, 0.7, 0), cex.lab=1.2, las=1)
legend("topleft",inset=c(0.22,0.02),legend=c("Observations"), col=grey(0.4), lty=NA, pch=20, lwd=1, cex=1, xpd=TRUE, bty="n", x.intersp=0.1, text.font=1)
```

## HSGP modeling and fitting

Input vector

```{r message=FALSE, warn=FALSE}
x <- scale(data$id[], center=TRUE, scale=TRUE)
```

Output vector

```{r message=FALSE, warn=FALSE}
y <- scale(data$births[], center=TRUE, scale=TRUE)
```

Standard deviation and mean of the output

```{r message=FALSE, warn=FALSE}
std_y <- attr(y,"scaled:scale")
m_y <- attr(y,"scaled:center")
```

Standard deviation of the input

```{r message=FALSE, warn=FALSE}
std_x <- attr(x,"scaled:scale")
```

Year and week periods

```{r message=FALSE, warn=FALSE}
period_year <- 365.25/std_x
period_week <- 7/std_x
```

`Day of the year2` is a variable that indicates the correspondence between all the days with the days of a leap-year

```{r message=FALSE, warn=FALSE}
day_of_year2 <- data$day_of_year2
```

### HSGP tunnig factors

Number of basis functions $M$ for the smooth trend function $f_1$

```{r message=FALSE, warn=FALSE}
M_f1 <- 40
```

Boundary factor $c$ for $f_1$

```{r message=FALSE, warn=FALSE}
c_f1 <- 1.5
```

Number of basis functions $J$ for year effects function $f_2$

```{r message=FALSE, warn=FALSE}
J_f2 <- 20		
```

Number of basis functions $J$ for the week effects function $f_3$

```{r message=FALSE, warn=FALSE}
J_f3 <- 7
```

### Data to be passed to Stan

```{r message=FALSE, warn=FALSE}
standata <- list(M_f1= M_f1, 
                 c_f1= c_f1, 
                 J_f2= J_f2, 
                 J_f3= J_f3, 
                 x= x[,1], 
                 y= y[,1], 
                 N= length(x), 
                 period_year= period_year, 
                 period_week= period_week,
                 day_of_year2= day_of_year2
)
```

### Stan model

```
//saved in 'stancode_def.stan'
functions {
  vector diagSPD_EQ(real gpscale, real lscale, real L, int M) {
    return sqrt((gpscale^2) * sqrt(2*pi()) * lscale * exp(-0.5*(lscale*pi()/2/L)^2 * linspaced_vector(M, 1, M)^2));
  }
  vector diagSPD_periodic(real gpscale, real lscale, int M) {
    real a = 1/lscale^2;
    int one_to_M[M];
    for (m in 1:M) one_to_M[m] = m;
    vector[M] q = sqrt(gpscale^2 * 2 / exp(a) * to_vector(modified_bessel_first_kind(one_to_M, a)));
    return append_row(q,q);
  }
  matrix PHI_EQ(int N, int M, real L, vector x) {
    return sin(diag_post_multiply(rep_matrix(pi()/(2*L) * (x+L), M), linspaced_vector(M, 1, M)))/sqrt(L);
  }
  matrix PHI_periodic(int N, int M, real w0, vector x) {
    matrix[N,M] mw0x = diag_post_multiply(rep_matrix(w0*x, M), linspaced_vector(M, 1, M));
    return append_col(cos(mw0x), sin(mw0x));
  }
}
data {
    real c_f1;                  //boundary value for function 1
    int<lower=1> M_f1;          //num basis functions for function 1
    int<lower=-1> J_f2;         //num cosine and sinu functions for function 3
    int<lower=-1> J_f3;         //num cosine and sinu functions for function 4
    int<lower=1> N;             //num observations
    vector[N] x;                //input vector
    vector[N] y;                //target vector
    real period_year;           //period of the year
    real period_week;           //period of the week
    int day_of_year2[N];        //day of the year inside a leap-year
}
transformed data {
    real L_f1= c_f1*max(fabs(x));
    //Basis functions for f1, f2 and f3
    matrix[N,M_f1] PHI_f1 = PHI_EQ(N, M_f1, L_f1, x);
    matrix[N,2*J_f2] PHI_f2 = PHI_periodic(N, J_f2, 2*pi()/period_year, x);
    matrix[N,2*J_f3] PHI_f3 = PHI_periodic(N, J_f3, 2*pi()/period_week, x);
}
parameters {
    vector[N] f;
    real intercept;
    //variables for the basis function models
    vector[M_f1] beta_f1;
    vector[2*J_f2] beta_f2;
    vector[2*J_f3] beta_f3;
    //hyperparameters
    vector<lower=0>[3] lscale;
    vector<lower=0>[3] gpscale;
    real<lower=0> noise;
    //
    //t-student prior for special days effects
    vector[366] f4;
    real<lower=0> sigma_f4;
}
transformed parameters{
    vector[N] f1;
    vector[N] f2;
    vector[N] f3;
    {
    vector[M_f1] diagSPD_f1 = diagSPD_EQ(gpscale[1], lscale[1], L_f1, M_f1);
    vector[2*J_f2] diagSPD_f2 = diagSPD_periodic(gpscale[2], lscale[2], J_f2);
    vector[2*J_f3] diagSPD_f3 = diagSPD_periodic(gpscale[3], lscale[3], J_f3);
    //
    vector[M_f1] SPD_beta_f1 = diagSPD_f1 .* beta_f1;
    vector[2*J_f2] SPD_beta_f2 = diagSPD_f2 .* beta_f2;
    vector[2*J_f3] SPD_beta_f3 = diagSPD_f3 .* beta_f3;
    //
    f1 = PHI_f1[,] * SPD_beta_f1;
    f2 = PHI_f2[,] * SPD_beta_f2;
    f3 = PHI_f3[,] * SPD_beta_f3;
    }
}
model{
    intercept ~ normal(0,1);
    beta_f1 ~ normal(0,1);
    beta_f2 ~ normal(0,1);
    beta_f3 ~ normal(0,1);
    //
    lscale ~ normal(0,2);           //GP lengthscales
    gpscale ~ normal(0,10);         //GP magnitudes
    noise ~ normal(0,1);            //model noise
    //
    //t-student prior for special days effects
    f4 ~ student_t(1, 0, sigma_f4);
    sigma_f4 ~ normal(0, 0.1);
    //
    f= intercept + f1 + f2 + f3 + f4[day_of_year2];
    target += normal_lpdf(y | f, noise);
}
generated quantities{
  vector[N] y_rep;
  vector[N] log_lik;
  for(n in 1:N){
    y_rep[n] = normal_rng(f[n], noise);
    log_lik[n] = normal_lpdf(y[n] | f[n], noise);
  }
}
```

### Model sampling

Compiling the model

```{r message=FALSE, warn=TRUE, eval=TRUE}
birthday_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_def.stan")
```

Model sampling (using `cmdstanr` package)

```{r warn=TRUE, message=FALSE, eval=TRUE}
fit <- birthday_mod$sample(data= standata, iter_warmup=200, iter_sampling=200, chains=4, thin=4, init=0.5, adapt_delta=0.9, save_warmup=FALSE)
```

<!-- Save or load model output -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
# save(fit, file="fit.rData")
# load("fit.rData")
```

<!-- Permanent saving of the output CSV files in the local directory -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
fit$save_output_files(dir = getwd())
```

Conversion of the model output from 'cmdstanr' package to 'rstan' package. This uses the output CSV files saved at a temporary directory. If you want a permament saving of these files at a local directory use `cmdstanr::fit$save_output_files()`

```{r message=FALSE, warn=FALSE, eval=TRUE}
fit_rstan <- rstan::read_stan_csv(fit$output_files())
```

<!-- Diagnostics of the sampling -->

```{r message=TRUE, warn=TRUE, eval=FALSE, include=FALSE}
check_hmc_diagnostics(fit_rstan)
```

## Model diagnosis

Posterior samples

```{r message=FALSE, warn=FALSE, eval=TRUE}
posterior <- as.array(fit_rstan)
```

Log posterior density

```{r message=FALSE, warn=FALSE, eval=TRUE}
lp <- log_posterior(fit_rstan)
```

Values for parameters of the performance of NUTS sampler

```{r message=FALSE, warn=FALSE, eval=TRUE}
np <- nuts_params(fit_rstan)
```

Draws

```{r message=FALSE, warn=FALSE, eval=TRUE}
draws <- extract(fit_rstan)
```

Summaries of variable estimates

```{r message=FALSE, warn=FALSE, eval=TRUE}
param = c("intercept","lscale[1]","lscale[2]","lscale[3]","gpscale[1]","gpscale[2]","gpscale[3]","noise","sigma_f4")
summary(fit_rstan, pars = param, probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
```

Simulation chains for the variables after warmup

```{r message=FALSE, warn=FALSE, fig.height=10, fig.width=15, eval=TRUE}
traceplot(fit_rstan, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

Bivariate samples plots

```{r message=FALSE, warn=FALSE, fig.height=10, fig.width=15, eval=TRUE}
mcmc_pairs(posterior, np= np, pars= param, off_diag_args= list(size = 1))
```

Autocorrelation plots

```{r message=FALSE, warn=FALSE, fig.height=5, fig.width=8, eval=TRUE}
mcmc_acf(posterior, pars = param, lags = 10)
```

Posterior predictive checks

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_dens_overlay(y[,1], draws$y_rep)
```

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_stat(y[,1], draws$y_rep, stat = "median")
```

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ppc_stat(y[,1], draws$y_rep, stat = "mean")
```

Leave-one-out PIT values using `loo` package

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
loo1 <- loo(fit_rstan, save_psis = TRUE, cores = 2)
psis1 <- loo1$psis_object
lw <- weights(psis1)
ppc_loo_pit_overlay(y[,1], draws$y_rep, lw)
```

Number of effective samples ratios for variables

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
ratios <- neff_ratio(fit_rstan, pars= param)
mcmc_neff(ratios, size = 2) + yaxis_text(hjust = 0)
```

Rhat statistics for sampling chains

```{r message=FALSE, warn=FALSE, fig.height=2, fig.width=4, eval=TRUE}
rhats <- rhat(fit_rstan, pars= param)
mcmc_rhat(rhats) + yaxis_text(hjust = 0)
```

Divergence

```{r message=FALSE, warn=FALSE, fig.height=4, fig.width=4, eval=TRUE}
mcmc_nuts_divergence(np, lp)
```

Energy

```{r message=FALSE, warn=FALSE, fig.height=5, fig.width=6, eval=TRUE}
mcmc_nuts_energy(np)
```

## Ploting estimated functions

Extracting estimated function components $f_1$, $f_2$, $f_3$ and $f_4$

```{r message=FALSE, warn=FALSE, eval=TRUE}
f <- summary(fit_rstan, pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f1 <- summary(fit_rstan, pars = c("f1"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f2 <- summary(fit_rstan, pars = c("f2"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f3 <- summary(fit_rstan, pars = c("f3"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
f4 <- summary(fit_rstan, pars = c("f4"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary
```

#### Plot of only one year, the year 1972

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_year <- data[data$year==1972,]
ind <- data_year$id
axis_labels_at <- aggregate(data_year, by=list(data_year$month), FUN=min)$id

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=21, bg=grey(0.7), cex=0.9, col=grey(0.4), xlab="", ylab="", lwd=1, ylim=c(0.73,1.16), mgp= c(2, 1, 0), frame.plot = TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")
axis(1, at=axis_labels_at, labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"), tick=TRUE, lty=1, mgp=c(2, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab ="Month", mgp= c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab ="Proportion of births over the mean", mgp= c(3, 0.7, 0), cex.lab=1.6, las=1)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col=grey(0.6), lwd=1)	# f
abline(h=1, lty=2)						                              # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)			    # f1 smooth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)	 		    # f3 year effect
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=2) 		    # f5 horseshoe

#labels special days
text(ind[data_year$month==1&data_year$day==1]+2, y = 0.84, labels = "New year", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==2&data_year$day==14], y = 1.045, labels = "Valentine's day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==2&data_year$day==29], y = 0.978, labels = "Leap day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==4&data_year$day==1]+2, y = 0.980, labels = "April 1st", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==5&data_year$day==27], y = 0.98, labels = "Memorial day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==7&data_year$day==4], y = 0.86, labels = "Independence day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==9&data_year$day==2], y = 0.97, labels = "Labor day", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==10&data_year$day==30]+2, y = 0.985, labels = "Halloween", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==11&data_year$day==25]+5, y = 0.95, labels = "Thanks-giving", pos =NULL, offset = 0, family="serif", cex=1.3)
text(ind[data_year$month==12&data_year$day==25], y = 0.82, labels = "Christmas", pos =NULL, offset = 0, family="serif", cex=1.3)

legend("topleft", inset=c(0.05,0.01), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Special days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,6,grey(0.5)), lty=c(NA,1,1,1,1), pch=c(20,NA,NA,NA,NA), lwd=c(2,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```

#### Plot of all the years

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
ind <- data$id
labels_at = aggregate(data, by=list(data$year), FUN=min)$id
plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=20, bg=grey(0.4), cex=0.6, col=grey(0.5), xlab="", ylab="", lwd=1, ylim=c(0.7,1.3), mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n", fg=grey(0.5), family="serif")

axis(1, at=labels_at, c("1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988"), tick=TRUE, lty=1, mgp=c(3, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Year", mgp=c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab="Proportion of births over the mean", mgp=c(3, 0.7, 0), cex.lab=1.6, las=1)

abline(h=1, lty=2)												                # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)				# smooth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)				# year effect

legend("topleft", inset=c(0.22,0.02), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)')), col=c(grey(0.4),2,3,"grey"), lty=c(NA,1,1,1), pch=c(20,NA,NA,NA), lwd=c(1,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=1, seg.len=1.5)
```

#### Plot of only the first month

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_month <- data[data$month==1&data$year==1972,]
ind <- data_month$id
axis_labels_at <- aggregate(data_month, by=list(data_month$day), FUN=min)$id

id_week <- data_month$id[data_month$day_of_week==1]

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", pch=21, bg=grey(0.7), cex=1.2, col=grey(0.4), xlab="", ylab="", lwd=1, ylim=c(0.7,1.2), mgp=c(3.5, 1, 0), frame.plot=TRUE, yaxs="r", cex.axis=1.6, cex.lab=1.6, las=1, xaxt="n", yaxt="n",fg=grey(0.5), family="serif")

axis(1, at=axis_labels_at[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31)], labels=as.character(1:31)[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29,31)], tick=TRUE, lty=1, mgp=c(3, 1, 0), las=1, cex.axis=1.6, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(1, at=id_week, labels=rep(c("Monday"),5), tick=TRUE, lty=1, mgp=c(-1, -1.2, 0), las=1, cex.axis=1.3, font=1, col=grey(0.5), col.ticks=grey(0.3), family="")
axis(2, at=NULL, labels=TRUE, tick=TRUE, lty=1, mgp=c(3, 0.7, 0), las=1, cex.axis=1.6, font=5, col=grey(0.5), col.ticks=grey(0.3))
title(xlab="Day", mgp=c(3, 1, 0), cex.lab=1.6, las=1)
title(ylab="Proportion of births over the mean", mgp=c(3, 0.7, 0), cex.lab=1.6, las=1)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col="black", lwd=2)	# f
abline(h=1, lty=2)											                  # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=2)			  # f1 smoth trend
lines(ind, (f2[ind,1]*std_y+m_y)/m_y, col=3, lwd=2)	 		  # f3 year effect
lines(ind, (f3[ind,1]*std_y+m_y)/m_y, col=4, lwd=2)	 		  # f4 week effect
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=2) 		  # f5 horseshoe

abline(v=id_week, lty=2, col="grey")

#labels special days
text(ind[data_month$month==1&data_month$day==1]-0.05, y = 0.84, labels = "New year", pos =NULL, offset = 0, family="serif", cex=1.3, xpd=TRUE,)

legend("topleft", inset=c(0.107,0.005), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Week effects ($f_3$)'),TeX('Special-days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,4,6,"black"), lty=c(NA,1,1,1,1,1), pch=c(20,NA,NA,NA,NA,NA), lwd=c(3,3,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```

#### Plot of the first four years

```{r message=FALSE, warn=FALSE, fig.height=8, fig.width=14, eval=TRUE}
data_year <- data[data$year==1969 | data$year==1970 | data$year==1971 | data$year==1972,]
ind <- data_year$id
axis_labels_at <- aggregate(data_year, by=list(data_year$year), FUN=min)$id

plot(ind, (y[ind]*std_y+m_y)/m_y, type="p", lty=1, pch=18, cex=0.4, col="black", xlab="", ylab="Proportion of births over the mean", cex.lab=1.5, cex.axis=1.6, xaxt="n", ylim=c(0.7,1.3))

axis(1, at=axis_labels_at, labels=c("1969","1970","1971","1972"), tick=TRUE, cex.axis=1.6)

lines(ind, (f[ind,1]*std_y+m_y)/m_y, col="grey", lwd=1)		# f
lines(range(ind), c(1,1), lty=2, lwd=2)						        # mean
lines(ind, (f1[ind,1]*std_y+m_y)/m_y, col=2, lwd=1)			  # f1
lines(ind, (f3[ind,1]*std_y+m_y)/m_y, col=3, lwd=1)	 		  # f3
lines(ind, (f4[data$day_of_year2,1][ind]*std_y+m_y)/m_y, col=6, lwd=1) 		  # f5

legend("topleft", inset=c(0.05,0.005), legend=c("Observations",TeX('Long-term trend ($f_1$)'),TeX('Year effects ($f_2$)'),TeX('Week effects ($f_3$)'),TeX('Special-days effects ($f_4$)'),TeX('$\\mu$=f_1+f_2+f_3+f_4$')), col=c(grey(0.5),2,3,4,6,"black"), lty=c(NA,1,1,1,1,1), pch=c(20,NA,NA,NA,NA,NA), lwd=c(3,3,3,3,3,3), cex=1.4, xpd=TRUE, bty="n", y.intersp=1, x.intersp=0.8, text.font=1, ncol=2, seg.len=1.3)
```