### 2022-05-15 Paper/Case-study_2D&3D-Simulated-data/2Ddataset/Rnb_case2D_diagnosis.Rmd
reorganize files

---
title: '2D simulated case study: Diagnosis of the HSGP approximation'
author: 'Gabriel Riutort-Mayol'
date: "`r format(Sys.Date())`."
output:
  bookdown::html_document2:
    theme: readable
    toc: yes
    toc_depth: 4
    number_sections: true
    toc_float: 
      collapsed: true
    code_download: yes
    df_print: kable
---

<style type="text/css">
body, td{ font-size: 14px; }
code.r{ font-size: 12px; }
pre{ font-size: 12px }
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
}
h3 { /* Header 3 */
    font-size: 14px;
  color: Black;
}
h4 { /* Header 4 */
    font-size: 14px;
  color: Black;
}
#TOC {
  color: DarkBlue;
  font-size: 16px;
}
</style>

This notebook implements the HSGP model fit and diagnosis on the 2D dataset in case study 5.3 '*Simulated data for 2D and 3D functions*' in [Riutort-Mayol, G., Bürkner, P. C., Andersen, M. R., Solin, A., & Vehtari, A. (2020; updated in 2022). Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming. arXiv preprint arXiv:2004.11408.](https://arxiv.org/pdf/2004.11408.pdf)

```{r include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment="")
```

# Load packages

```{r message=FALSE, warn=FALSE, eval=TRUE}
library(rstan)
library(cmdstanr)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = FALSE)
library(posterior)
library(bayesplot)
library(RColorBrewer)
library(ggplot2)
```

# Synthetic dataset

A synthetic dataset $\boldsymbol{y}$ with a true data generating function $f(\boldsymbol{x})$, with $\boldsymbol{x}_i \in {\rm I\!R}^2$, where $f(\boldsymbol{x})$ is a Gaussian process (GP) with a squared exponential kernel. Gaussian errors $\boldsymbol{\epsilon}$ are added to $f(\boldsymbol{x})$ to form the final synthetic noisy dataset $\boldsymbol{y}$.

\begin{align*}
\begin{split}
\boldsymbol{y} &= \boldsymbol{f} + \boldsymbol{\epsilon} \\
\boldsymbol{\epsilon} &\sim \text{Normal}(\boldsymbol{0}, \sigma^2 \boldsymbol{I}) \\
f(\boldsymbol{x}) &\sim \text{GP}(0, k(\boldsymbol{x}, \boldsymbol{x}', \theta)),
\end{split}
\end{align*}

where $k(\boldsymbol{x}, \boldsymbol{x}', \theta)$ is a squared exponential kernel as a function of input values $\boldsymbol{x}$ and parameters $\theta=(\alpha,\boldsymbol{\ell})$, where $\alpha$ is the marginal variance of the GP function and $\boldsymbol{\ell}=(\ell_1,\ell_2)$ the lengthscale of the kernel for dimensions 1 and 2, respectively.

## Draw the true data generative function $f(x)$ from a GP

### Input predictive space

Matrix of 2D-predictive input points `x_pred`

```{r eval=TRUE}
x_pred <- expand.grid(x1= seq(-1,1,by=0.04), x2= seq(-1,1,by=0.04))  
str(x_pred)
```

Length of `x_pred` (`N_pred`)

```{r eval=TRUE}
N_pred <- dim(x_pred)[1]
N_pred
```

Number of dimensions `D`

```{r eval=TRUE}
D <- 2 
D
```

Half range `S` of every input dimension

```{r eval=TRUE}
S1 <- abs((max(x_pred$x1) - min(x_pred$x1))/2)
S1
S2 <- abs((max(x_pred$x2) - min(x_pred$x2))/2)
S2
```

### Draw $f(x)$ using Stan

Marginal variance `gpscale` ($\alpha$) and multivariate lengthscale `lscale` ($\boldsymbol{\ell}$)

```{r eval=TRUE}
gpscale <- 1.1
lscale <- vector()
lscale[1] <- 0.1
lscale[2] <- 0.3
```

Random variable `eta` to simulate from the GP

```{r eval=TRUE}
seed <- 88646
set.seed(seed)
eta <- rnorm(N_pred, 0, 1)
```

Data to be passed to Stan

```{r eval=TRUE}
standata <- list(D= D, 
                 N_pred= N_pred, 
                 x_pred= x_pred,
                 gpscale= gpscale, 
                 lscale= lscale, 
                 eta= eta)
```

Model sampling

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
simuGP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_SIMU_2D.stan")
fit_SIMU <- simuGP_mod$sample(data= standata,
                                iter_warmup=1,
                                iter_sampling=1,
                                chains=1, thin=1, init=0.5,
                                fixed_param=TRUE)
fit_SIMU <- read_stan_csv(fit_SIMU$output_files())
```

Drawn $f(x)$

```{r eval=TRUE}
f_true <- extract(fit_SIMU, pars="f_true", permuted = TRUE, inc_warmup = FALSE, include = TRUE)$f_true[1,]
```

```{r eval=TRUE, echo=FALSE}
xx <- unique(x_pred[,1])

col1 <- brewer.pal(11,"BrBG")[11:1] 
brk <- seq(-4.00,max(f_true), length.out= 12)

filled.contour(xx, xx, matrix(f_true,length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03))
text(-0.3,0.9,"(a)", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

## Adding noise to $f(x)$

```{r eval=TRUE}
set.seed(1234)
y_pred <- f_true + rnorm(N_pred, 0, 0.2)
```

## Sampling data set

```{r eval=TRUE}
nobs <- 150 
set.seed(4567)
vv_sample <- sample(1:N_pred, nobs)
N_sample <- length(vv_sample)
```

# GP model fitting

## Data

Data to Stan

```{r eval=TRUE}
standata_GP <- list(D= D, 
                    N_sample= N_sample,
                    vv_sample= vv_sample,
                    N_pred= N_pred, 
                    x_pred= x_pred, 
                    y_pred= y_pred) 
str(standata_GP)
```

## Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
GP_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_GP_marginalized_2D.stan")
fit_GP <- GP_mod$sample(data= standata_GP,
                          iter_warmup=50,
                          iter_sampling=50,
                          chains=4, thin=1, init=2,
                          save_warmup=FALSE)
fit_GP <- read_stan_csv(fit_GP$output_files())
```

Summaries and simulation chains for the variables after warmup

```{r fig.height=5, fig.width=10, eval=TRUE}
param = c("lscale","gpscale[1]","noise")
print(summary(fit_GP, pars = param, probs = c(0.025, 0.5, 0.975))$summary)
traceplot(fit_GP, pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

## Posterior predictive distribution

```{r eval=TRUE}
f_gp <- summary(fit_GP, pars = c("f_pred"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]
```

```{r eval=TRUE, echo=FALSE}
col1 <- brewer.pal(11,"BrBG")[11:1] 
# brk <- seq(min(f_gp[,1]),max(f_gp[,1]), length.out= 12)

filled.contour(xx, xx, matrix(f_gp[,1],length(xx),length(xx)), levels = brk, col = brewer.pal(11,"BrBG")[11:1], plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.7)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.7), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03)) #; grid(col=grey(0.90), lty=1)
text(-0.3,0.9,"(b)   GP", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

## Error plot

```{r eval=TRUE}
error <- f_true - f_gp[,1]
```

```{r eval=TRUE, echo=FALSE}
col3 <- c(colorRampPalette(c("blue", "white"), bias=10)(6), colorRampPalette(c("white", "red"), bias=10)(6)[2:6])#brewer.pal(11,"RdBu")[1:11]
brk_error <- seq(min(error),max(error), length.out= 12)

filled.contour(xx, xx, matrix(error,length(xx),length(xx)), levels= brk_error, col= col3, plot.axes= {axis(1, tick=FALSE, line=-1.1, font=5, mgp= c(2.5, 0.1, 0), las=1, cex.axis=1.5); title(xlab =expression(x[1]), mgp= c(0.8, 1, 0), cex.lab=2, las=1, family="serif"); axis(2, tick=FALSE, line=0.4, font=5, mgp= c(2.5, 0.0, 0), las=1, cex.axis=1.5); points(x_pred[vv_sample,], cex=1, lwd=1, pch=1);  abline(h= seq(-1,1,0.5), col=grey(0.6)); lines(c(-1,-1,-0.5,-0.5,0,0,0.5,0.5,1,1), c(-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03,-1.03,1.03), col=grey(0.6), type="h")}, asp= 1, frame.plot=FALSE, xlim=c(-1.03,1.03), ylim=c(-1.03,1.03))
text(-0.3,1.12,"(a)", font=2, cex=1.5);
text( par("usr")[1] - 0.30, mean(c(-1,1)), labels = expression(x[2]), srt = 0, xpd = TRUE, cex=2, font=1, family="serif")
```

# Functional relationships among $m$, $\ell$ and $c$ parameters in the HSGP approach

Relationship among the number of basis functions $m$, the lengthscale $\ell$ and the boundary factor $c$ for a **squared exponential kernel** in the HSGP approach:

$$\begin{align*}
m= 1.75 \frac{c}{\frac{\ell}{S}} \;\; \Leftrightarrow \;\; \frac{\ell}{S}= 1.75 \, \frac{c}{m}  (\#eq:m-l-QE)
\end{align*}$$

```{r eval=TRUE}
m_QE <- function(c,l,S) ceiling(1.75 * c / (l/S))
l_QE <- function(c,m,S) round(S * 1.75 * c / m, 3)   
```

with

$$\begin{align*}
c \, \geq \, 3.2 \, \frac{\ell}{S}\;\; \& \;\;c \, \geq \frac{1}{2}  (\#eq:c-QE)
\end{align*}$$

as a function of $\ell$ and $S$, where $S$ is the half range of the input domain

```{r eval=TRUE}
c_vs_l_QE <- function(l,S){
    c =  3.2*l/S
    if(c < 1.2)
      c = 1.2
  c
}
```

# Lengthscale diagnostic

The HSGP lengthscale diagnostic is defined as:

$$\begin{align*}
\hat{\ell} + 0.01 \geq \ell  (\#eq:diagnostic)
\end{align*}$$

```{r eval=TRUE}
diagnostic <- function(l,l_hat) l_hat + 0.01 >= l
```

# HSGP fit and diagnosis

Initializing programming objects

```{r eval=TRUE}
standata <- list()
fit<- list()
diagnosis <- list()

l1 <- vector()  # lengthscale for dimension 1
l2 <- vector()  # lengthscale for dimension 2
c1 <- vector()  # boundary factor for dimension 1
c2 <- vector()  # boundary factor for dimension 2
m1 <- vector()  # number of basis functions for dimension 1
m2 <- vector()  # number of basis functions for dimension 2
l_hat1 <- vector()  # HSGP lengthscale estimate for dimension 1
l_hat2 <- vector()  # HSGP lengthscale estimate for dimension 2
check1 <- vector()  # Diagnostic for dimension 1
check2 <- vector()  # Diagnostic for dimension 2
rmse_gp <- vector()  # root mean square error of GP model
rmse <- vector()  # root mean square error of HSGP model
eR <- vector()  # coefficient of determination of HSGP model
elpd <- vector()  # expected log predictive density of HSGP model
```

## Iteration 1

```{r eval=TRUE}
i <- 1
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If iteration $i=1$:

1. Making a first guess for lengthscale $\ell_i$

2. The boundary factor $c_i$ has to fulfill equation \@ref(eq:c-QE)

3. The number of basis functions $m_i$ has to fulfill equation \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(i==1){
  l1[i] <- 0.5
  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)

  l2[i] <- 1
  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
		mm = mm+1
		indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * c(S1, S2), 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f_hsgp[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_gp[vv_sample] - f_hsgp[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell}_i$ in the HSGP approximation

```{r eval=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell}_i + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 2

```{r eval=TRUE}
i <- 2
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i=2$:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i==2){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
}

if(diagnostic(l2[i-1], l_hat2[i-1]) & i==2){
  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)
}

if(!diagnostic(l2[i-1], l_hat2[i-1])){
  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * c(S1, S2), 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f_hsgp[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_gp[vv_sample] - f_hsgp[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 3)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 3)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 3

```{r eval=TRUE}
i <- 3
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is FALSE:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
}

if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & !diagnostic(l2[i-2], l_hat2[i-2])){  
  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1}$

2. Update $c_i = c_{i-1}$

3. Update $\ell_i = \ell_{i-1}$

```{r eval=TRUE}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1]
  c1[i] <- c1[i-1]
  l1[i] <- l1[i-1]
}

if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & diagnostic(l2[i-2], l_hat2[i-2])){  
  m2[i] <- m2[i-1]
  c2[i] <- c2[i-1]
  l2[i] <- l2[i-1]
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)
}

if(!diagnostic(l2[i-1], l_hat2[i-1])){
  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * c(S1, S2), 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f_hsgp[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_gp[vv_sample] - f_hsgp[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$ \@ref(eq:diagnostic)

```{r eval=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```

## Iteration 4

```{r eval=TRUE}
i <- 4
print(paste("Iteration i =", i), quote = FALSE)
```

### Setting $m$, $\ell$ and $c$

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} \geq \ell_{i-2}$ \@ref(eq:diagnostic) is FALSE:

1. Update $m_i = m_{i-1} + 2$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE) where $c(\cdot)$ is evaluated as a function of $\hat{\ell}_{i-1}$

3. The lengthscale $\ell_i$ has to fulfill \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & !diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1] + 2
  c1[i] <- c_vs_l_QE(l=l_hat1[i-1], S=S1)
  l1[i] <- l_QE(c=c1[i], m=m1[i], S=S1)
}

if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & !diagnostic(l2[i-2], l_hat2[i-2])){ 
  m2[i] <- m2[i-1] + 2
  c2[i] <- c_vs_l_QE(l=l_hat2[i-1], S=S2)
  l2[i] <- l_QE(c=c2[i], m=m2[i], S=S2)
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is TRUE and iteration $i>2$ and diagnostic $\hat{\ell}_{i-2} + 0.01 \geq \ell_{i-2}$ \@ref(eq:diagnostic) is TRUE:

1. Update $m_i = m_{i-1}$

2. Update $c_i = c_{i-1}$

3. Update $\ell_i = \ell_{i-1}$

```{r eval=TRUE}
if(diagnostic(l1[i-1], l_hat1[i-1]) & i>2 & diagnostic(l1[i-2], l_hat1[i-2])){
  m1[i] <- m1[i-1]
  c1[i] <- c1[i-1]
  l1[i] <- l1[i-1]
}

if(diagnostic(l2[i-1], l_hat2[i-1]) & i>2 & diagnostic(l2[i-2], l_hat2[i-2])){ 
  m2[i] <- m2[i-1]
  c2[i] <- c2[i-1]
  l2[i] <- l2[i-1]
}
```

If the diagnostic $\hat{\ell}_{i-1} + 0.01 \geq \ell_{i-1}$ \@ref(eq:diagnostic) is FALSE:

1. Update $\ell_i = \hat{\ell}_{i-1}$

2. The boundary factor $c_i$ has to fulfill \@ref(eq:c-QE)

3. The number of basis functions $m_i$ as a function of $\ell_i$ and $c_i$ by using \@ref(eq:m-l-QE)

```{r eval=TRUE}
if(!diagnostic(l1[i-1], l_hat1[i-1])){
  l1[i] <- l_hat1[i-1]
  c1[i] <- c_vs_l_QE(l=l1[i], S=S1)
  m1[i] <- m_QE(c=c1[i], l=l1[i], S=S1)
}

if(!diagnostic(l2[i-1], l_hat2[i-1])){
  l2[i] <- l_hat2[i-1]
  c2[i] <- c_vs_l_QE(l=l2[i], S=S2)
  m2[i] <- m_QE(c=c2[i], l=l2[i], S=S2)
}
```

### Matrix of $D$-tuples

Matrix of $D$-tuples as in equation (10) in the [paper](https://arxiv.org/pdf/2004.11408.pdf)

```{r eval=TRUE}
indices <- matrix(NA, m1[i]*m2[i], D)
mm=0;
for (r in 1:m1[i]){
	for (s in 1:m2[i]){
			mm = mm+1
			indices[mm,] = c(r, s)
	}
}
```

### Data to be passed to Stan

```{r eval=TRUE}
standata[[i]] <- list(D= D, 
                      M= c(m1[i], m2[i]),
                      M_nD= m1[i]*m2[i], 
                      indices= indices,
                      L= c(c1[i], c2[i]) * c(S1, S2), 
                      x_pred= x_pred, 
                      y_pred= y_pred, 
                      N_sample= N_sample, 
                      N_pred= N_pred, 
                      vv_sample= vv_sample
)
str(standata[[i]])
```

### Model fitting

Model fitting

```{r eval=TRUE, results=FALSE, message=FALSE, warn=FALSE}
hsgp_mod <- cmdstanr::cmdstan_model(stan_file = "stancode_HSGP_2D.stan")
fit[[i]] <- hsgp_mod$sample(data= standata[[i]],
                                iter_warmup=100,
                                iter_sampling=100,
                                chains=4, thin=2, init=0.5,
                                # adapt_delta=0.9,
                                save_warmup=TRUE)
fit[[i]] <- rstan::read_stan_csv(fit[[i]]$output_files())
```

Summaries and simulation chains of variable estimates

```{r eval=TRUE}
param = c("lscale[1]","lscale[2]","gpscale","noise")
summary(fit[[i]], pars = param, probs = c(0.025, 0.5, 0.975))$summary
traceplot(fit[[i]], pars = param, include = TRUE, unconstrain = FALSE, inc_warmup = FALSE, window = NULL, nrow = NULL, ncol = NULL)
```

### Model evaluation

Root mean square error `rmse`, Bayesian $R^2$ (Coefficient of determination) `eR`, and expected log predictive density `elpd`.

```{r eval=TRUE}
f_hsgp <- summary(fit[[i]], pars = c("f"), probs = c(0.025, 0.5, 0.975), digits_summary = 4)$summary[,c(1,4,6)]

# rmse against data
residual <- standata[[i]]$y_pred[vv_sample] - f_hsgp[,1]
rmse[i] <- round(sqrt(mean(residual^2)), 2)

# rmse against gp
residual_gp <- f_gp[vv_sample] - f_hsgp[,1]
rmse_gp[i] <- round(sqrt(mean(residual_gp^2)), 2)

# Coefficient of determination R^2
noise <- as.matrix(fit[[i]], pars = c("noise"))
sd_f <- apply(as.matrix(fit[[i]], pars = c("f")), 1, sd)
eR[i] <- round(median(sd_f^2/(sd_f^2 + noise^2)), 2)

# elpd
elpd[i] <- round(median(summary(fit[[i]], pars = c("elpd"))$summary[,1]), 2)
```

### Diagnosis

Lengthscale estimate $\hat{\ell_i}$ in the HSGP approximation

```{r eval=TRUE}
l_hat1[i] <- round(summary(fit[[i]], pars = "lscale[1]", probs = c(0.05, 0.95))$summary[,1], 2)
l_hat2[i] <- round(summary(fit[[i]], pars = "lscale[2]", probs = c(0.05, 0.95))$summary[,1], 2)
```

Check the diagnostic of whether $\, \hat{\ell_i} + 0.01 \geq \ell_i$

```{r eval=TRUE}
check1[i] <- diagnostic(l1[i], l_hat1[i])
check2[i] <- diagnostic(l2[i], l_hat2[i])
```

Summary table of diagnosis

  - `rmse_gp` refers to the root mean square error of the HSGP function with the GP function as the reference.

  - `rmse` refers to the root mean square error of the HSGP function with the true function as the reference.

```{r eval=TRUE}
diagnosis[[i]] <- data.frame(iter= c(i, i),
                              Dim= c("D1", "D2"),
                              l= c(l1[i], l2[i]),
                              c= c(c1[i], c2[i]),
                              m= c(m1[i], m2[i]),
                              l_hat= c(l_hat1[i], l_hat2[i]),
                              'l_hat > l' = c(check1[i], check2[i]),
                              rsme_gp = c(NA, rmse_gp[i]),
                              rsme = c(NA, rmse[i]),
                              R2 = c(NA, eR[i]),
                              elpd = c(NA, elpd[i]))
names(diagnosis[[i]]) <- c("iter", "Dim", "l", "c", "m", "l_hat", "l_hat + 0.01 > l", "rmse_gp", "rmse", "R^2", "elpd")

if(i==1){ 
  diagnosis[[i]]
}else{
  diagnosis[[i]] <- rbind(diagnosis[[i-1]],diagnosis[[i]])
  diagnosis[[i]]
}
```